### 2018-08-14 FernandesWeaverBertoldi_AdvMaterials_2018.tex
added all paper files

\documentclass[9pt,twocolumn,twoside]{fernandes_paper} 
%Default opts:9pt,twocolumn,twoside
\graphicspath{{images_paper/}}

\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}

\title{Utilizing Design Principles from Glass Sponges for Structurally Robust Lattices}

\author[1]{Matheus C. Fernandes}
\author[2]{James C. Weaver}
\author[1,3,*]{Katia Bertoldi}

\affil[1]{John A. Paulson School of Engineering and Applied Sciences -- Harvard University, Cambridge, MA 02138}
\affil[2]{Wyss Institute -- Harvard University, Cambridge, MA 02138}
\affil[3]{Kavli Institute -- Harvard University, Cambridge, MA 02138}
\affil[*]{Corresponding author: \href{mailto:bertoldi@seas.harvard.edu}{bertoldi@seas.harvard.edu}}

\keywords{Architected materials, truss structure, buckling resistance, glass sponges, bio-inspired engineering.}

\usepackage{comment}

\begin{document}
\maketitle
% \doublespace
%\linenumbers

\section{Introduction}
\subsection{Sponge Biology}
\subsection{History of Diagonally Reinforced Structures and Previous Works}
\subsection{Examples of Current Diagonally Reinforced Designs}

\section{Methods}
\subsection{Numerical Model Setup}
\subsection{Experimental Design}

\section{Results}
\subsection{Linear Stiffness for different Designs}
 \subsection{Buckling Resistance for Different Designs}
\subsection{Experimental Results}

\section{Discussion}
\subsection{Parameter Exploration}
\subsection{Parameter Optimization}

\section{Conclusion}

\section*{Acknowledgments}
\section*{References}

\begin{figure*}[ht]
  \captionsetup{width=0.8\textwidth}
\begin{center}
\includegraphics[width=0.9\textwidth]{Fig1}
\end{center}
\caption{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge. (c) Close up microscope image of the sponge. (d) Comparison between the idealized model (green and blue lines) and the sponge structure. (e) Representative volume element unit cell of the idealized model.}\label{Fig1}
\end{figure*}

\begin{figure*}[ht]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.9\textwidth]{Fig2}
	\caption{text}\label{Fig2}
\end{figure*}

\begin{figure*}[ht]
	\captionsetup{width=0.8\textwidth}
	\begin{center}
		\includegraphics[width=0.7\textwidth]{Fig3}
	\end{center}
	\caption{\textbf{text} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs. The designs considered are depicted below as Design A-D.} \label{Fig3}
\end{figure*}

% \section{Introduction}
The foundational structure of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}. In modern day structures, we see many different iterations of new truss designs with various applications in mind but with what is the same fundamentally simple diagonal reinforcement technique. This technique, involving two diagonal trusses (one in each direction) crossing the opening of a square beam cell, limits the transverse motion of the main support beams ultimately giving the structure shear capacity and strength. Alternative, methods to strengthen beam structures have been introduced via material \hl{[...],} \mf{needs more information and citations}. However, in this study we seek to further strengthen square lattices through inspiration from a particular sponge species called \textit{Euplectella Aspergillum}, commonly known as the Venus' flower basket. This particular species, which is part of the \textit{Hexactinellid} class, is primarily made of silica and exhibit a periodic and regular grid structure with an intrinsic and different diagonal reinforcement scheme.

Hexactinellids, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m. Beyond their fracture propagation inhibiting material composition \citep{weaver2007}, they are perceived to exhibit large structural rigidity and strength against buckling. 
% Since they are primarily made of 'brittle silica', buckling strength may be a crucial property in making them resistant to impact and environmentally applied stresses. 
Structurally, they are composed of a base square-grid architecture and regular ordering of vertical and horizontal struts that form the skeletal system. Furthermore, their base structure is overlaid with double diagonal (two diagonals in each direction) reinforcement struts, which create a checkerboard-like pattern of open-closed cell structure. This diagonal reinforcement design is conjectured to give the sponge greater buckling resistance and strength to localized damage then it would experience having a single diagonal reinforcement strut (with equivalent allocation of mass between the diagonal reinforcements.) 
% Analogous to the sponge, many engineering structures, such as buildings and bridges, exhibit diagonal reinforcement struts as a stability mechanism. Based on this similarity, we explore the following research question:
The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?} \mf{this paragraph needs more citations and more avid lit. survey}

In this paper, we present a numerical analysis of the structure deformation under various loading conditions as well as survey different arrangements within similar design space of the sponge via parametric studies. Through the various design iterations we look for the critical buckling strain and the elastic load carrying capacity. Finally, we compare the results obtained from the sponge design to what is typically used in engineering structures and validate our results with experimental data.

% involving the development of truss structures with efficient diagonal reinforcement, is inspired by glass sponges (primarily made of silica) commonly known as the Venus Flower Basket. These deep sea sponges exhibit a regular and periodic truss structure which biologists for many years have stipulated to be an optimized design for structural and hydrodynamic purposes. To investigate these claims, I have built numerical and analytical models analyzing the buckling effects of the sponge designs by tuning various geometrical parameters and analyzing its effect on the buckling response. Thus far, we have found the the double diagonal reinforced sponge design delays the onset of structural buckling by a factor of twice the applied strain in comparison to the traditional designs. We believe that this intricate approach inspired by the sponge can improve the way we construct buildings and bridges, which will give the opportunity to build taller buildings and bridges that use less material and are stronger.

% \section{Results}
% \subsection{Linear Stiffness for different Designs}
In this section, we compare the linear stiffness for the different arrangements using Finite Element Analysis (FEA). For each design, we consider a Representative Volume Element (RVE) and apply periodic boundary conditions along the edges of the unit cell domains depicted along the bottom of \cref{Fig3}. In our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections. In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.

The results presented in \cref{Fig3}(a) show that all of the structures containing diagonal reinforcement behave the same way for varying loading angles. This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to the perpendicular aligned beam elements. This is further supported by the graph for Design D, where all of the material from the diagonal beams is removed and allocated to the non-diagonal elements, giving the structure a higher stiffness at $0^\circ$ and lower stiffness at $45^\circ$. Any stiffness at $45^\circ$ in Design D is strictly a result of the bending resistance due to the pinned joints between the non-diagonal beams.

% \subsection{Buckling Resistance for Different Designs}
The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements. From this figure, we can see that Design A, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement. At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of Design C, which is the commonly used truss systems.

% \subsection{Experimental Validation}
For the experimental validation, we conducted various uniaxial compression tests of 3D printed planar beam structures (with circular cross-section) on an Instron 5566. The specimen was 3D printed using \hl{xxxx} material with manufacturer reported \hl{xxxx GPa} modulus. The printed specimen consisted of a 10x10 cell structure spanning a total of 10x10 inch planar structure. The experimental set-up consisted of two \hl{1 inch} thickness acrylic sheets bolted together along the perimeter and spaced from each other by the thickness of the 3D printed specimen.\\\\\hl{[...]} \mf{more to go here when we have experimental setup and data}

% \figll{Buckling_Separation}{0.9}{Critical buckling strain for varying diagonal separation in Design A parametric sweep. The normalized separation distance $C/L$ indicates how far apart the diagonals are from each other. For $C/L=0$ we obtain the Design B and for $C/L=0.5$ we obtain the diagonals crossing half way through the length of the non-diagonal trusses ($L$). For all values in this sweep, we assume a mass ratio equivalent to that of the sponge. The dashed gray line indicates the assumed diagonal separation of the sponge of $C/L=1/(\sqrt{2}+2)$.}

In order to study the impact of diagonal separation for Design A (sponge design), we perform a parametric study on the buckling behavior of the structure under varying separation distances ($C$). Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles. \Cref{Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$). Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.

% \section{Mass Ratio Parametrization}
% \figll{Stiffness_MassRatio}{0.9}{Structural stiffness for varying mass ratio parametric sweep. Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements. For all values in this graph, the design is kept constant while the cross-sectional area of each element is scaled accordingly. For all values of $\lambda$, the total mass of all elements are kept constant. The different colored lines correspond to different loading angles of the unit cell. The gray dashed line indicates the measured mass ratio for the sponge. For all values of this sweep, we assume Design A geometry although the linear stiffness does not depend on the design.}

% \figll{Buckling_MassRatio}{0.9}{Critical buckling strain for varying mass ratio parametric sweep. Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements. For all values in this graph, the design is kept constant while the cross-sectional area of each element is scaled accordingly. For all values of $\lambda$, the total mass of all elements are kept constant. The different colored lines correspond to different loading angles of the unit cell. The gray dashed line indicates the measured mass ratio for the sponge. For all values of this sweep, we assume Design A geometry.}

% \section{Diagonal Separation Parametrization}

To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements. To quantitatively describe this mass ratio, we introduce the variable $\lambda$ as the fraction $V_{nd}/V_d$, where $V_{nd}$ describes the collective volume of all non-diagonal elements and $V_d$ is the collective volume of all diagonal elements in the unit cell. It is important to note, that for all simulations varying $\lambda$ we maintain the same total mass of the structure, namely $V_{nd}+V_d=\mbox{const. } \forall (V_{nd},V_d)$.

\Cref{Fig5}(b) shows the structural stiffness as a function of $\lambda$ for various loading angles (corresponding to the respective colors). In this figure, we see that for small $\lambda$ the structure is stiffer at $0^\circ$ loading and for large $\lambda$ the structure is stiffer at $45^\circ$. This is expected as we are allocating more material to its respective elements. Furthermore, we can see that at $\lambda=1$ we attain a point in the mass ratio where the stiffness is equal for all angles. That is because there is the same amount of material allocated to the diagonals as there is to the non-diagonal elements.

\Cref{Fig5}(c) illustrates the critical buckling strain as a function of $\lambda$. Here, we can clearly see that buckling strain is dependent on the mass ratio and behaves differently for all angles \hl{[...]} \mf{I am not sure what to say here (interpretation) about this graph.}

% \section{Conclusion}
% \hl{add writing}

\section*{Funding Information}
\hl{add writing}
% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship

\section*{Acknowledgments}
\hl{add writing} 
% Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft

%\begin{figure*}
%	\captionsetup{width=0.8\textwidth}
%	\begin{center}
%		\includegraphics[width=0.7\textwidth]{Fig5}
%	\end{center}
%	\caption{\textbf{Parametric sweep results for different analysis.} (a) Critical buckling strain for varying diagonal separation in Design A. The normalized separation distance $C/L$ indicates how far apart the diagonals are from each other. For all values we assume a mass ratio to be equivalent to that of the sponge. The dashed gray line indicates the assumed diagonal separation of the sponge, namely $C/L=1/(\sqrt{2}+2)$. (b) Structural linear stiffness for varying mass ratio. Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements based on the volume element. For all values of $\lambda$, the total mass of all elements are kept constant. The gray dashed line indicates the measured mass ratio for the sponge. (c) Critical buckling strain for varying mass ratio. For all values of $\lambda$, the total mass of all elements are kept constant. The different colored lines correspond to different loading angles of the unit cell. The gray dashed line indicates the measured mass ratio for the sponge. For all values in (b) and (c), we assume Design A geometry.} \label{Fig5}
%\end{figure*}

 Bibliography
\nocite{*}
\bibliography{refs}
\bibliographystyle{apalike}

\end{document}

### 2018-08-14 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
added all paper files

\documentclass[10pt,twoside]{fernandes_supp} 
%Default opts:9pt,twocolumn,twoside
\graphicspath{{images_supp/}}

\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
\newcommand{\MF}[1]{\noindent\color{ForestGreen}#1\normalcolor}

\title{Supplemental Information:\\ Harnessing  Design Principles from Glass Sponges for Structurally Robust Lattice Structures}

\author[1]{Matheus C. Fernandes}
\author[2]{James C. Weaver}
\author[1,3,*]{Katia Bertoldi}

\affil[1]{John A. Paulson School of Engineering and Applied Sciences -- Harvard University, Cambridge, MA 02138}
\affil[2]{Wyss Institute -- Harvard University, Cambridge, MA 02138}
\affil[3]{Kavli Institute -- Harvard University, Cambridge, MA 02138}
\affil[*]{Corresponding author: \href{mailto:bertoldi@seas.harvard.edu}{bertoldi@seas.harvard.edu}}

\begin{document}
\maketitle
% \doublespace
% \linenumbers
\section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
The periodic structures investigated in this study are inspired by the skeleton of the hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket". In this section we provide a detailed description of the sponge geometry and measured dimensions.

\Cref{Sponge} shows a photograph of the entire skeleton of the \textit{Euplectella sp.}, and its intricate, cylindrical cage-like structure (20 to 25cm long, 2 to 4 cm in diameter) \citep{aizenberg2005}. The surface of the cylinder consists of a regular square lattice composed of a series of cemented vertical and horizontal struts with circular cross-section. The cell spacing between horizontal and vertical struts was reported  to be $L\approx 2.5$mm \citep{weaver2007}, while their diameter was measured to be $D_{nd}\approx 0.25$mm \citep{weaver2007}. Besides the horizontal and vertical struts, there is an additional set of diagonal elements, intersecting in a manner that creates a series of alternating open and closed cells, reminiscent of a checkerboard pattern \citep{weaver2007}. Although these diagonal elements are not as ordered as the horizontal and vertical ones, it has been shown that they can be approximated with two diagonal struts that are offset from the nodes  (vertex joints between non-diagonal elements) and form octagonal openings (\cref{Sponge}(d)). To estimate the volume ratio between diagonal and non-diagonal elements, we took high resolution photographs of the sponge and performed image segmentation to segregate the projected area of the vertical/horizontal and diagonal spicules. Using this approach, the projected area ratio of non-diagonal to diagonal elements was found to be $A_{nd}/A_{d}\approx1.4$. Note that here, and in the following, the subscripts $d$ and $nd$ are used to indicate diagonal and non-diagonal (i.e. horizontal and vertical) elements, respectively.

Finally, it should also be noted that the sponge is reinforced by external ridges that extend perpendicular to the surface of the cylinder and spiral the cage at an angle of $45^\text{o}$. However, in this paper we do not report the effects of these ridges on it's structural performance %and treat the periodic structure as a planar (no out of plane deformation) unit cell.
% \figll{SpongeCylinder}{0.35}{Close-up photograph of Hexactinellid sponge Euplectella aspergillum.}
\figll{Sponge}{0.9}{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge.  (c)  Close up microscope image of the sponge. (d) Comparison between the idealized model (green and blue lines) and the sponge structure. (e) Unit cell of the idealized model.}

\section{Designs Considered}
In this study, we consider four different lattice configurations using in two different elements cross-sections constrained to deform in an in-plane setting only. In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations the total amount of material (i.e. volume or mass) is conserved. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry. In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

%In this study, we consider four different planar periodic lattice structures and compare their performance in terms of buckling resistance and stiffness. All structures are made of struts with circular cross section and are assumed to only deform in-plane.  In an effort to provide a fair performance comparison between all four designs, we scale the different diameters of each design to maintain a constant total volume for each unit cell. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry. In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

\subsection{Polygonal Cross-section}
\subsubsection{Design A}
From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ . Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
\begin{equation}
	V_{A,nd}=8LD_{A,nd}t
\end{equation}
\begin{equation}
	V_{A,d}=8\sqrt{2}LD_{A,d}t
\end{equation}
Let us enter this into our biological observation, which yields:
\begin{equation}
	{V_{A,nd}\over V_{A,d}}\approx 1.4\approx\sqrt{2}={8LD_{A,nd}t\over8\sqrt{2}LD_{A,d}t}
\end{equation}
Therefore, we can simplify this relationship as
\begin{equation}
	D_{A,nd}=2D_{A,d}
\end{equation}

\subsubsection{Design B}
For this design the volume of non-diagonal elements remain the same, namely,
\begin{equation}
		V_{B,nd}=8LD_{B,nd}t.
\end{equation}
However, the volume for the diagonal elements will is different as a result the change in total length, namely,
\begin{equation}
		V_{B,d}=4\sqrt{2}LD_{A,d}t.
\end{equation}
In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
\begin{equation}
		{V_{B,nd}\over V_{B,d}}\approx 1.4\approx\sqrt{2}={8LD_{B,nd}t\over4\sqrt{2}LD_{B,d}t}
\end{equation}
which simplifies to
\begin{equation}
	D_{B,nd}=D_{B,d}
\end{equation}
and
\begin{equation}
	D_{B,nd}=D_{A,nd}
\end{equation}

\subsubsection{Design C}
Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
\begin{equation}
D_{A,nd}=2D_{A,d}
\end{equation}
and
\begin{equation}
D_{C,nd}=D_{A,nd}
\end{equation}

\subsubsection{Design D}
For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements. Thus we can formulate the total volume of Design A as
\begin{equation}
	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
\end{equation}
Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
\begin{equation}
V_{A,nd}+V_{A,d}=\left(1+{1\over\sqrt{2}}\right)8LtD_{A,nd}
\end{equation}

We can build a relationship between Design A and Design D through the total volume, such that the total volume of Design A is the same as the total volume of Design D
\begin{equation}
\left(1+{1\over\sqrt{2}}\right)8LtD_{A,nd}=8LtD_{D,nd}
\end{equation}
which yields a relationship between the thicknesses of the non diagonal elements as
\begin{equation}
	D_{D,nd}=\left(1+{1\over\sqrt{2}}\right)D_{A,nd}
\end{equation}

\subsection{Cylindrical Cross-section}

\subsubsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{TwoDiag}). If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
respectively. Note that in this study we use 
\begin{equation}
\frac{D_{A,nd}}{L}=0.1,
\end{equation}
since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

Moreover, as for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell. As such, the diagonals intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes and their  volume and projected area in the unit cell are
\begin{equation}\label{V2}
V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
and
\begin{equation}\label{A2}
A_{A,d}=8\sqrt{2} L D_{A,d},
\end{equation}
respectively. Since the projected area ratio of the non-diagonal to diagonal elements in the sponge has been measured to be 
\begin{equation}
\frac{A_{A,nd}}{A_{A,d}}=1.4,
\end{equation}
by substituting \cref{A1} and \cref{A2} into the equation above we find that for Design A
\begin{equation} \label{DA}
D_{A,nd}=1.4\sqrt{2}D_{A,d}\approx 2 D_{A,d}.
\end{equation}
Substitution of \cref{DA} into \cref{V1} and \cref{V2} yields 
\begin{equation}
\frac{V_{A,nd}}{V_{A,d}}=\frac{2 L \pi D^2_{A,nd}}{2\sqrt{2}L \pi D^2_{A,d}}=2\sqrt{2}
\end{equation}
and
\begin{equation}\label{VT}
V_{A,T}=V_{A,nd}+V_{A,d}=2\pi L (D_{A,nd}^2+\sqrt{2} D_{A,d}^2)=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right),
\end{equation}
where $V_{A,T}$ indicates the total volume of the unit cell for Design A. 
% such that \cref{VT} can be further simplified to
% \begin{equation}
% {V}_{A,T}=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right)\approx1.3288[\text{mm}^3].
% \end{equation}

Finally, it is important to note that in this study we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs to be equal to that of Design A, namely,
\begin{equation}\label{con1}
V_{\alpha,d}+V_{\alpha,nd}={V}_{A,T}=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right),
\end{equation}
with $\alpha=$ B, C and D.
Finally,  for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A
\begin{equation}\label{con2}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\frac{V_{A,nd}}{V_{A,d}}=2\sqrt{2},
\end{equation}
with $\alpha=$ B and C.

\figll{TwoDiag}{0.45}{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system. The horizontal and vertical struts have length $L$ and circular cross section with diameter $D_{A,nd}$ and, as with the sponge, we assume $D_{A,nd}/L=0.1$. The diagonal elements have a circular cross section  with diameter $D_{A,d}=2 D_{A,nd}$.}

\subsubsection{Design B}
Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{OneDiag}). However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. It follows that the non-diagonal and diagonal volumes are given by
\begin{equation}
V_{B,nd}=V_{A,nd}=2\pi L D_{B,nd}^2
\end{equation}
and
\begin{equation}
V_{B,d}=2\sqrt{8} L \left(\pi \frac{{D}_{B,d}^2}{4}\right),
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2}, as well as the above volumes, we  obtain 
\begin{equation}
{{D}_{B,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{B,d}}{{D}_{B,nd}}=\frac{1}{\sqrt{2}}.
\end{equation}

 \figll{OneDiag}{0.45}{{\bf Unit cell for Design B.} This design is still characterized by an alternation of open and closed cells. However, instead of having two  diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{B,nd}$. The diagonal elements have a circular cross section  with diameter $D_{B,d}={D_{B,nd}}/{\sqrt{2}}$.}

\subsubsection{Design C}
Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{FullDiag}). For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
\begin{equation}
V_{C,nd}=V_{A,nd}=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}
V_{C,d}=V_{A,d}=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2} we  obtain 
\begin{equation}
{{D}_{C,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{C,d}}{{D}_{C,nd}}=\frac{1}{2}.
\end{equation}

\figll{FullDiag}{0.45}{{\bf Unit cell for Design C.} This design consists of a square grid with all cells being reinforced by diagonal trusses passing through the nodes.  The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{C,nd}$. The diagonal elements have a circular cross section  with diameter $D_{C,d}={D_{C,nd}}/{2}$.}

\subsubsection{Design D} 
Design D comprises only the square grid without diagonal reinforcement (\cref{NoDiag}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}. As such, for this design we allocate the total material volume to the non-diagonal elements. Since
\begin{equation}
V_{D,T}=V_{D,nd}=V_{A,nd}=2\pi L D_{D,nd}^2,
\end{equation}
using the constraint provided by \cref{con1} we obtain
\begin{equation}
{D_{D,nd}}={D_{A,nd}}\sqrt{1+\frac{\sqrt{2}}{4}}. 
\end{equation}

\figll{NoDiag}{0.45}{{\bf Unit cell for Design D.} This design consists of a square grid without diagonal reinforcement.  The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{D,nd}$.}

%\subsection{Analytical Derivation Of Stiffness}
%\subsection{Analytical Derivation of Critical Buckling Strain}
%\subsection{Derivation for }

% Bibliography
\nocite{aizenberg2005}
\nocite{deshpande2001}
\nocite{miserez2008}
\nocite{weaver2010}

\bibliography{refs}
\bibliographystyle{apalike}
% \bibliographystyle{plainnat}

\end{document}

### 2018-08-23 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
Updates from ShareLaTeX

--- 2018-08-14 FernandesWeaverBertoldi_AdvMaterials_2018.tex
@@
 % \doublespace
-%\linenumbers
+\linenumbers
 
-\section{Introduction}
-\subsection{Sponge Biology}
-\subsection{History of Diagonally Reinforced Structures and Previous Works}
-\subsection{Examples of Current Diagonally Reinforced Designs}
+% \section{Introduction}
+% \subsection{Sponge Biology}
+% \subsection{History of Diagonally Reinforced Structures and Previous Works}
+% \subsection{Examples of Current Diagonally Reinforced Designs}
 
-\section{Methods}
-\subsection{Numerical Model Setup}
-\subsection{Experimental Design}
+% \section{Methods}
+% \subsection{Numerical Model Setup}
+% \subsection{Experimental Design}
 
-\section{Results}
-\subsection{Linear Stiffness for different Designs}
- \subsection{Buckling Resistance for Different Designs}
-\subsection{Experimental Results}
+% \section{Results}
+% \subsection{Linear Stiffness for different Designs}
+%  \subsection{Buckling Resistance for Different Designs}
+% \subsection{Experimental Results}
 
-\section{Discussion}
-\subsection{Parameter Exploration}
-\subsection{Parameter Optimization}
+% \section{Discussion}
+% \subsection{Parameter Exploration}
+% \subsection{Parameter Optimization}
 
-\section{Conclusion}
+% \section{Conclusion}
 
-\section*{Acknowledgments}
-\section*{References}
+% \section*{Acknowledgments}
+% \section*{References}
 
@@
 % \section{Introduction}
-The foundational structure of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
+
+In modern day structural engineering, we find many
+
+The design of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
 In modern day structures, we see many different iterations of new truss designs with various applications in mind but with what is the same fundamentally simple diagonal reinforcement technique.
@@
 Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles.
-\Cref{Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$).
+{Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$).
 Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.
@@
 
-\Cref{Fig5}(b) shows the structural stiffness as a function of $\lambda$ for various loading angles (corresponding to the respective colors).
+{Fig5}(b) shows the structural stiffness as a function of $\lambda$ for various loading angles (corresponding to the respective colors).
 In this figure, we see that for small $\lambda$ the structure is stiffer at $0^\circ$ loading and for large $\lambda$ the structure is stiffer at $45^\circ$.
@@
 
-\Cref{Fig5}(c) illustrates the critical buckling strain as a function of $\lambda$.
+{Fig5}(c) illustrates the critical buckling strain as a function of $\lambda$.
 Here, we can clearly see that buckling strain is dependent on the mass ratio and behaves differently for all angles \hl{[...]} \mf{I am not sure what to say here (interpretation) about this graph.}

### 2018-09-05 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
Updates from ShareLaTeX
Before Katia begins to edit

@@
 \newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
-\newcommand{\MF}[1]{\noindent\color{ForestGreen}#1\normalcolor}
+\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}
 
@@
 \section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
-The periodic structures investigated in this study are inspired by the skeleton of the hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket".
+The periodic structures investigated in this study are inspired by the skeleton of the Hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket".
 In this section we provide a detailed description of the sponge geometry and measured dimensions.
@@
 Finally, it should also be noted that the sponge is reinforced by external ridges that extend perpendicular to the surface of the cylinder and spiral the cage at an angle of $45^\text{o}$.
-However, in this paper we do not report the effects of these ridges on it's structural performance %and treat the periodic structure as a planar (no out of plane deformation) unit cell.
-% \figll{SpongeCylinder}{0.35}{Close-up photograph of Hexactinellid sponge Euplectella aspergillum.}
-\figll{Sponge}{0.9}{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge.
+However, in this paper we do not report the effects of these ridges on it's structural performance.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\linewidth]{SFig1.png}
+    \caption{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge.
 (c)  Close up microscope image of the sponge.
@@
 (e) Unit cell of the idealized model.}
+    \label{Sponge}
+\end{figure}
 
@@
 In this study, we consider four different lattice configurations using in two different elements cross-sections constrained to deform in an in-plane setting only.
-In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations the total amount of material (i.e.
+In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations, the total amount of material (i.e.
 volume or mass) is conserved.
@@
 
-%In this study, we consider four different planar periodic lattice structures and compare their performance in terms of buckling resistance and stiffness.
-All structures are made of struts with circular cross section and are assumed to only deform in-plane.
-In an effort to provide a fair performance comparison between all four designs, we scale the different diameters of each design to maintain a constant total volume for each unit cell.
-Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry.
-In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).
-
-\subsection{Polygonal Cross-section}
-\subsubsection{Design A}
-From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ .
-Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
-\begin{equation}
-	V_{A,nd}=8LD_{A,nd}t
-\end{equation}
-\begin{equation}
-	V_{A,d}=8\sqrt{2}LD_{A,d}t
-\end{equation}
-Let us enter this into our biological observation, which yields:
-\begin{equation}
-	{V_{A,nd}\over V_{A,d}}\approx 1.4\approx\sqrt{2}={8LD_{A,nd}t\over8\sqrt{2}LD_{A,d}t}
-\end{equation}
-Therefore, we can simplify this relationship as
-\begin{equation}
-	D_{A,nd}=2D_{A,d}
-\end{equation}
-
-\subsubsection{Design B}
-For this design the volume of non-diagonal elements remain the same, namely,
-\begin{equation}
-		V_{B,nd}=8LD_{B,nd}t.
-\end{equation}
-However, the volume for the diagonal elements will is different as a result the change in total length, namely,
-\begin{equation}
-		V_{B,d}=4\sqrt{2}LD_{A,d}t.
-\end{equation}
-In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
-\begin{equation}
-		{V_{B,nd}\over V_{B,d}}\approx 1.4\approx\sqrt{2}={8LD_{B,nd}t\over4\sqrt{2}LD_{B,d}t}
-\end{equation}
-which simplifies to
-\begin{equation}
-	D_{B,nd}=D_{B,d}
-\end{equation}
-and
-\begin{equation}
-	D_{B,nd}=D_{A,nd}
-\end{equation}
-
-\subsubsection{Design C}
-Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
-\begin{equation}
-D_{A,nd}=2D_{A,d}
-\end{equation}
-and
-\begin{equation}
-D_{C,nd}=D_{A,nd}
-\end{equation}
-
-\subsubsection{Design D}
-For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements.
-Thus we can formulate the total volume of Design A as
-\begin{equation}
-	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
-\end{equation}
-Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
-\begin{equation}
-V_{A,nd}+V_{A,d}=\left(1+{1\over\sqrt{2}}\right)8LtD_{A,nd}
-\end{equation}
-
-We can build a relationship between Design A and Design D through the total volume, such that the total volume of Design A is the same as the total volume of Design D
-\begin{equation}
-\left(1+{1\over\sqrt{2}}\right)8LtD_{A,nd}=8LtD_{D,nd}
-\end{equation}
-which yields a relationship between the thicknesses of the non diagonal elements as
-\begin{equation}
-	D_{D,nd}=\left(1+{1\over\sqrt{2}}\right)D_{A,nd}
-\end{equation}
-
 \subsection{Cylindrical Cross-section}
+For this cross-section type we stay consistent with the geometry found within the biological species.
+Although, this geometry is difficult to realize and validate experimentally, the result for this geometry pose great value to the understanding of the biological species.
 
 \subsubsection{Design A}
-Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{TwoDiag}).
+Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}).
 If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by
@@
 
-\figll{TwoDiag}{0.45}{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\linewidth]{SFig2.png}
+    \caption{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system.
 The horizontal and vertical struts have length $L$ and circular cross section with diameter $D_{A,nd}$ and, as with the sponge, we assume $D_{A,nd}/L=0.1$.
 The diagonal elements have a circular cross section  with diameter $D_{A,d}=2 D_{A,nd}$.}
+    \label{DesignA}
+\end{figure}
 
 \subsubsection{Design B}
-Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{OneDiag}).
+Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{DesignB}).
 However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell.
@@
 
- \figll{OneDiag}{0.45}{{\bf Unit cell for Design B.} This design is still characterized by an alternation of open and closed cells.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\linewidth]{SFig3.png}
+    \caption{{\bf Unit cell for Design B.} This design is still characterized by an alternation of open and closed cells.
 However, instead of having two  diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell.
@@
 The diagonal elements have a circular cross section  with diameter $D_{B,d}={D_{B,nd}}/{\sqrt{2}}$.}
+    \label{DesignB}
+\end{figure}
 
 \subsubsection{Design C}
-Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{FullDiag}).
+Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{DesignC}).
 For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
@@
 
-\figll{FullDiag}{0.45}{{\bf Unit cell for Design C.} This design consists of a square grid with all cells being reinforced by diagonal trusses passing through the nodes.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\linewidth]{SFig4.png}
+    \caption{{\bf Unit cell for Design C.} This design consists of a square grid with all cells being reinforced by diagonal trusses passing through the nodes.
 The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{C,nd}$.
 The diagonal elements have a circular cross section  with diameter $D_{C,d}={D_{C,nd}}/{2}$.}
+    \label{DesignC}
+\end{figure}
 
 \subsubsection{Design D} 
-Design D comprises only the square grid without diagonal reinforcement (\cref{NoDiag}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}.
+Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}.
 As such, for this design we allocate the total material volume to the non-diagonal elements.
@@
 
-\figll{NoDiag}{0.45}{{\bf Unit cell for Design D.} This design consists of a square grid without diagonal reinforcement.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\linewidth]{SFig5.png}
+    \caption{{\bf Unit cell for Design D.} This design consists of a square grid without diagonal reinforcement.
 The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{D,nd}$.}
+    \label{DesignD}
+\end{figure}
+
+\subsection{Cylindrical Cross-section Results}
+\mf{add results and figures here}
+Include graphs showing results for cylindrical cross-section, as can be found in the main paper.
+
+\subsection{Polygonal Cross-section}
+In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section.
+For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior.
+The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
+
+To the determine the minimum cross-section depth a parametric simulation on the necessary depth was run for each structure.
+The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.
+
+\subsubsection{Design A}
+From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ .
+Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
+\begin{equation}
+	V_{A,nd}=8LD_{A,nd}t
+\end{equation}
+\begin{equation}
+	V_{A,d}=8\sqrt{2}LD_{A,d}t
+\end{equation}
+Let us enter this into our biological observation, which yields:
+\begin{equation}
+	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
+\end{equation}
+Therefore, we can simplify this relationship as
+\begin{equation}
+	D_{A,nd}=2D_{A,d}
+\end{equation}
+
+\subsubsection{Design B}
+For this design the volume of non-diagonal elements remain the same, namely,
+\begin{equation}
+		V_{B,nd}=8LD_{B,nd}t.
+\end{equation}
+However, the volume for the diagonal elements will is different as a result the change in total length, namely,
+\begin{equation}
+		V_{B,d}=4\sqrt{2}LD_{A,d}t.
+\end{equation}
+In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
+\begin{equation}
+		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
+\end{equation}
+which simplifies to
+\begin{equation}
+	D_{B,nd}=D_{B,d}
+\end{equation}
+and
+\begin{equation}
+	D_{B,nd}=D_{A,nd}
+\end{equation}
+
+\subsubsection{Design C}
+Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
+\begin{equation}
+D_{A,nd}=2D_{A,d}
+\end{equation}
+and
+\begin{equation}
+D_{C,nd}=D_{A,nd}
+\end{equation}
+
+\subsubsection{Design D}
+For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements.
+Thus we can formulate the total volume of Design A as
+\begin{equation}
+	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
+\end{equation}
+Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
+\begin{equation}
+V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}
+\end{equation}
+
+We can build a relationship between Design A and Design D through the total volume, such that the total volume of Design A is the same as the total volume of Design D
+\begin{equation}
+\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}=8LtD_{D,nd}
+\end{equation}
+which yields a relationship between the thicknesses of the non diagonal elements as
+\begin{equation}
+	D_{D,nd}=\left(1+\frac{1}{\sqrt{2}}\right)D_{A,nd}
+\end{equation}
+
+\section{Domain Size Analysis}
+In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
+This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.
+
+\mf{Write this section}
+
+\section{Optimization Analysis Description}
+For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.
+The algorithm input variables are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction.
+Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements.
+For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+
+For the optimization results presented in the main article, the critical buckling load is the target response maximized.
+However, a an equivalent analysis is performed to maximize the critical buckling strain as the target response.
+This analysis is presented in \cref{BucklingOptimization}.
+Similarly, the same analysis was performed on the elastic stiffness of the structure and yields a trivial result leading to maximizing the allocation of mass to the non-diagonal elements through the mass ratio.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\linewidth]{SFig6.png}
+    \caption{Caption}
+    \label{BucklingOptimization}
+\end{figure}
+
+\section{Parameter Exploration}
+\mf{Write this section}
+Include the graphs showing critical buckling strain as a function of parameters.
 

### 2018-09-11 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
progress

@@
 \section{Designs Considered}
-In this study, we consider four different lattice configurations using in two different elements cross-sections constrained to deform in an in-plane setting only.
+In this study, we focus on four different lattice configurations constrained to deform in an in-plane setting only.
 In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations, the total amount of material (i.e.
 volume or mass) is conserved.
-Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry.
+Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry) and two different shapes for the cross section of all lattice members: circular and rectangular.
+For circular cross-sections  we stay consistent with the geometry found within the biological species.
+However, since slender members with circular cross section easily buckle out-of-plane when the structure is subjected to compression, to facilitate  fabrication and testing and to avoid out-of-plane buckling, we also consider a rectangular cross-section.
+\KB{how do you determine the aspect ratio of the rectangular cross section?
+I don't get the text that follows from here}
+
+For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior.
+The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
+
+In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section.
+For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior.
+The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
+
+To the determine the minimum cross-section depth a parametric simulation on the necessary depth was run for each structure.
+The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.
+
 In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).
 
-\subsection{Cylindrical Cross-section}
-For this cross-section type we stay consistent with the geometry found within the biological species.
+%For this cross-section type we stay consistent with the geometry found within the biological species.
 Although, this geometry is difficult to realize and validate experimentally, the result for this geometry pose great value to the understanding of the biological species.
 
-\subsubsection{Design A}
+\subsection{Design A}
 Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}). 
+\subsubsection{Circular cross section} 
 If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
@@
 where $V_{A,T}$ indicates the total volume of the unit cell for Design A.
-% such that \cref{VT} can be further simplified to
-% \begin{equation}
-% {V}_{A,T}=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right)\approx1.3288[\text{mm}^3].
-% \end{equation}
 
@@
 
-\subsubsection{Design B}
+\subsubsection{Rectangular cross-section}
+From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ .
+Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
+\begin{equation}
+	V_{A,nd}=8LD_{A,nd}t
+\end{equation}
+\begin{equation}
+	V_{A,d}=8\sqrt{2}LD_{A,d}t
+\end{equation}
+Let us enter this into our biological observation, which yields:
+\begin{equation}
+	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
+\end{equation}
+Therefore, we can simplify this relationship as
+\begin{equation}
+	D_{A,nd}=2D_{A,d}
+\end{equation}
+
+\subsection{Design B}
 Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{DesignB}).
 However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell.
+\subsubsection{Circular cross section}
 It follows that the non-diagonal and diagonal volumes are given by
@@
 
-\subsubsection{Design C}
+\subsubsection{Rectangular cross section}
+
+For this design the volume of non-diagonal elements remain the same, namely,
+\begin{equation}
+		V_{B,nd}=8LD_{B,nd}t.
+\end{equation}
+However, the volume for the diagonal elements will is different as a result the change in total length, namely,
+\begin{equation}
+		V_{B,d}=4\sqrt{2}LD_{A,d}t.
+\end{equation}
+In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
+\begin{equation}
+		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
+\end{equation}
+which simplifies to
+\begin{equation}
+	D_{B,nd}=D_{B,d}
+\end{equation}
+and
+\begin{equation}
+	D_{B,nd}=D_{A,nd}
+\end{equation}
+
+\subsection{Design C}
 Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{DesignC}).
+
+\subsubsection{Circular cross section}
 For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
@@
 
-\subsubsection{Design D} 
+\subsubsection{Rectangular cross section}
+Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
+\begin{equation}
+D_{A,nd}=2D_{A,d}
+\end{equation}
+and
+\begin{equation}
+D_{C,nd}=D_{A,nd}
+\end{equation}
+
+\subsection{Design D} 
 Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}.
 As such, for this design we allocate the total material volume to the non-diagonal elements.
+
+\subsubsection{Circular cross section}
 Since
@@
 
-\subsection{Cylindrical Cross-section Results}
-\mf{add results and figures here}
-Include graphs showing results for cylindrical cross-section, as can be found in the main paper.
-
-\subsection{Polygonal Cross-section}
-In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section.
-For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior.
-The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
-
-To the determine the minimum cross-section depth a parametric simulation on the necessary depth was run for each structure.
-The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.
-
-\subsubsection{Design A}
-From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ .
-Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
-\begin{equation}
-	V_{A,nd}=8LD_{A,nd}t
-\end{equation}
-\begin{equation}
-	V_{A,d}=8\sqrt{2}LD_{A,d}t
-\end{equation}
-Let us enter this into our biological observation, which yields:
-\begin{equation}
-	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
-\end{equation}
-Therefore, we can simplify this relationship as
-\begin{equation}
-	D_{A,nd}=2D_{A,d}
-\end{equation}
-
-\subsubsection{Design B}
-For this design the volume of non-diagonal elements remain the same, namely,
-\begin{equation}
-		V_{B,nd}=8LD_{B,nd}t.
-\end{equation}
-However, the volume for the diagonal elements will is different as a result the change in total length, namely,
-\begin{equation}
-		V_{B,d}=4\sqrt{2}LD_{A,d}t.
-\end{equation}
-In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
-\begin{equation}
-		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
-\end{equation}
-which simplifies to
-\begin{equation}
-	D_{B,nd}=D_{B,d}
-\end{equation}
-and
-\begin{equation}
-	D_{B,nd}=D_{A,nd}
-\end{equation}
-
-\subsubsection{Design C}
-Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
-\begin{equation}
-D_{A,nd}=2D_{A,d}
-\end{equation}
-and
-\begin{equation}
-D_{C,nd}=D_{A,nd}
-\end{equation}
-
-\subsubsection{Design D}
+\subsubsection{Rectangular cross-section}
 For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements.
@@
 
-\section{Domain Size Analysis}
+\section{Additional numerical results}
+\KB{add here all additional numerical results}
 In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
@@
 
-\section{Optimization Analysis Description}
+\section{Optimization Analysis}
+
+\KB{what target functions are you using?
+Which parameters are you identifying via the optimization?
+How do you choose the initial values for all variables?
+Which value do you use for the initial standard deviation?}
+
 For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 For circular cross-sections  we stay consistent with the geometry found within the biological species.
-However, since slender members with circular cross section easily buckle out-of-plane when the structure is subjected to compression, to facilitate  fabrication and testing and to avoid out-of-plane buckling, we also consider a rectangular cross-section.
-\KB{how do you determine the aspect ratio of the rectangular cross section?
-I don't get the text that follows from here}
+However, since slender members with circular cross section easily buckle out-of-plane when the structure is subjected to compression we must mitigate the issue by pursuing a plane-strain structure such that it's depth slenderness ratio is sufficiently deep to avoid out-of-plane buckling.
+In order to address this, we converted the biological observations of the cylindrical cross-section into a polygonal section.
+For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane deformation.
+The out-of-plane slenderness is scaled accordingly with mass to maintain the depth across all elements constant.
+To the determine the minimum cross-section depth a parametric simulation on depth was run to determine the necessary depth for each structure.
+The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.
 
@@
 The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
-
-In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section.
-For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior.
-The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
-
-To the determine the minimum cross-section depth a parametric simulation on the necessary depth was run for each structure.
-The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.
-
 In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).
-
-%For this cross-section type we stay consistent with the geometry found within the biological species.
-Although, this geometry is difficult to realize and validate experimentally, the result for this geometry pose great value to the understanding of the biological species.
 

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 The out-of-plane slenderness is scaled accordingly with mass to maintain the depth across all elements constant.
-To the determine the minimum cross-section depth a parametric simulation on depth was run to determine the necessary depth for each structure.
-The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.
+A parametric simulation on depth was analyzed to determine the minimum cross-section depth required to prevent out-of-plane buckling.
+Once the overall minimum depth for all geometries was chosen a factor of safety of 1.5 was added for the finial depth.
+The approach for minimum out-of-plane depth was further verified experimentally and the value was determined sufficient.
 
@@
 Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements.
-For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+For simulations with odd number of diagonal re-enforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
 
-For the optimization results presented in the main article, the critical buckling load is the target response maximized.
-However, a an equivalent analysis is performed to maximize the critical buckling strain as the target response.
+For the optimization results presented in the main article, we seek to maximize the singly objective target function for the critical buckling load.
+However, an equivalent analysis is performed to maximize the critical buckling strain as the target response.
 This analysis is presented in \cref{BucklingOptimization}.

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 
-\KB{what target functions are you using?
-Which parameters are you identifying via the optimization?
-How do you choose the initial values for all variables?
-Which value do you use for the initial standard deviation?}
-
 For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.
-The algorithm input variables are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction.
+The algorithm parameters (continuous input variables) are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction.
 Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements.
 For simulations with odd number of diagonal re-enforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and diagonal separation$=0.5*L$.
+The covariance matrix is initialized uniformly with standard deviation \mf{ADD HERE}.
 
-For the optimization results presented in the main article, we seek to maximize the singly objective target function for the critical buckling load.
+In optimization results presented in the main article, we seek to maximize the singly objective target function for the critical buckling load.
 However, an equivalent analysis is performed to maximize the critical buckling strain as the target response.

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 
-\section{Additional numerical results}
-\KB{add here all additional numerical results}
-In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
-This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.
-
 \mf{Write this section}
@@
 
+\section{Additional numerical results}
+In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
+This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.
+
 %\subsection{Analytical Derivation Of Stiffness}

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

--- 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
@@
 
-\section{Additional numerical results}
-\KB{add here all additional numerical results}
-In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
-This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.
-
-\mf{Write this section}
-
 \section{Optimization Analysis}
@@
 \section{Parameter Exploration}
+Include the graphs showing critical buckling strain as a function of parameters.\\
+\mf{Add figures for the diagonal}\\
+\mf{Add figure for $\lambda$ exploration}\\
 \mf{Write this section}
-Include the graphs showing critical buckling strain as a function of parameters.
+
+\section{Additional numerical results}
+\mf{Write this section}
+
+In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
+This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.
+
+\mf{Write about minimum RVE}
 

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 \section{Optimization Analysis}
-
 For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.
@@
     \includegraphics[width=0.9\linewidth]{SFig6.png}
-    \caption{Caption}
+    \caption{{\bf Buckling Optimization.} Graph shows optimal value of critical buckling strain for varying number of diagonals.
+For all simulations, the total mass of the structure is maintained constant while the mass-ratio is allowed to vary.
+Furthermore, the diagonal separation for each pair of diagonal is allowed to vary.
+}
     \label{BucklingOptimization}

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and diagonal separation$=0.5*L$.
+The optimization is run under a uniaxial loading condition in the direction parallel non-diagonal elements.
+In order to maintain reasonable computational complexity, the optimization was run on the unit cell constructed using one dimensional beams with node seeding of at least 1/20 of each beam length.
 The covariance matrix is initialized uniformly with standard deviation \mf{ADD HERE}.
@@
 For all simulations, the total mass of the structure is maintained constant while the mass-ratio is allowed to vary.
-Furthermore, the diagonal separation for each pair of diagonal is allowed to vary.
-}
+Furthermore, the diagonal separation for each pair of diagonal is allowed to vary together ensuring half symmetry of the structure at all times.
+The optimization is run under a uni-axial loading condition.}
     \label{BucklingOptimization}

### 2018-09-12 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 Furthermore, the diagonal separation for each pair of diagonal is allowed to vary together ensuring half symmetry of the structure at all times.
-The optimization is run under a uni-axial loading condition.}
+The optimization is run under a uniaxial loading condition.
+The color of each point on the graph corresponds to the optimum value of mass ratio for that particular number of diagonals.
+Four deformation plots of the optimum structure and buckling modes are provided on the right for their respective number of diagonals.}
     \label{BucklingOptimization}
@@
 
-\mf{Write about minimum RVE}
+\mf{Write about minimum RVE -- how do I convey this message? Should I produce a plot or just verbally state this check?}
 

### 2018-09-17 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.
progress

@@
 In order to maintain reasonable computational complexity, the optimization was run on the unit cell constructed using one dimensional beams with node seeding of at least 1/20 of each beam length.
-The covariance matrix is initialized uniformly with standard deviation \mf{ADD HERE}.
+The covariance matrix is initialized uniformly with standard deviation ($\sigma$) half of the domain space, which are normalized to remain between 0 and 1.
 
@@
 \mf{Add figure for $\lambda$ exploration}\\
+\mf{Make sure to include this analysis for both, rectangular cross-section and circular cross-section}\\
 \mf{Write this section}
+\subsection{Cylindrical Cross Section}
+\subsection{Rectangular Cross Section}
 
@@
 
-%\subsection{Analytical Derivation Of Stiffness}
+\section{Analytical Derivation Of Uniaxial Stiffness}
+\mf{Write this section}
 %\subsection{Analytical Derivation of Critical Buckling Strain}
-%\subsection{Derivation for }
 

### 2018-09-17 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 In this study, we focus on four different lattice configurations constrained to deform in an in-plane setting only.
-In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations, the total amount of material (i.e.
+In an effort to conduct a fair performance comparison between the different designs0, the total amount of material (i.e.
 volume or mass) is conserved.
@@
 In order to address this, we converted the biological observations of the cylindrical cross-section into a polygonal section.
-For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane deformation.
+For this new cross-section type, we assume the structure to be sufficiently deep in-plane in comparison the structural length scale (plane strain assumption) to avoid out-of-plane deformation.
+The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
 The out-of-plane slenderness is scaled accordingly with mass to maintain the depth across all elements constant.
@@
 
-For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior.
-The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
 In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

### 2018-09-18 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 
-\section{Designs Considered}
+\section{Designs Considered}\label{sec:designs}
 In this study, we focus on four different lattice configurations constrained to deform in an in-plane setting only.
-In an effort to conduct a fair performance comparison between the different designs0, the total amount of material (i.e.
-volume or mass) is conserved.
-Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry) and two different shapes for the cross section of all lattice members: circular and rectangular.
-For circular cross-sections  we stay consistent with the geometry found within the biological species.
-However, since slender members with circular cross section easily buckle out-of-plane when the structure is subjected to compression we must mitigate the issue by pursuing a plane-strain structure such that it's depth slenderness ratio is sufficiently deep to avoid out-of-plane buckling.
-In order to address this, we converted the biological observations of the cylindrical cross-section into a polygonal section.
-For this new cross-section type, we assume the structure to be sufficiently deep in-plane in comparison the structural length scale (plane strain assumption) to avoid out-of-plane deformation.
-The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.
-The out-of-plane slenderness is scaled accordingly with mass to maintain the depth across all elements constant.
-A parametric simulation on depth was analyzed to determine the minimum cross-section depth required to prevent out-of-plane buckling.
-Once the overall minimum depth for all geometries was chosen a factor of safety of 1.5 was added for the finial depth.
-The approach for minimum out-of-plane depth was further verified experimentally and the value was determined sufficient.
+In an effort to conduct a fair performance comparison between the different designs, the total volume of material is conserved.
+Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry) as well as two different shapes for the cross section of all lattice members: circular and rectangular.
+For the circular cross-section case  we stay consistent with the geometry found within the biological species.
+However, since slender members with such cross-section easily buckle out-of-plane when subject to compression loading we mitigate the issue by pursuing a plane-strain representative structure.
+To do so, we introduce the rectangular cross-section, where it's depth to slenderness ratio is sufficiently large to avoid out-of-plane deformation.
 
-In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).
+In the subsequent sections, we describe in detail the unit cells for four different designs (Designs A-D), and provide the derivations for each geometry cross-section values.
 
@@
 with $\alpha=$ B, C and D.
-Finally,  for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A
+For Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A
 \begin{equation}\label{con2}
@@
     \centering
-    \includegraphics[width=0.45\linewidth]{SFig2.png}
+    \includegraphics[width=0.4\linewidth]{SFig2.png}
     \caption{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system.
@@
 \subsubsection{Rectangular cross-section}
-From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ .
+From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$.
 Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
@@
     \centering
-    \includegraphics[width=0.45\linewidth]{SFig3.png}
+    \includegraphics[width=0.4\linewidth]{SFig3.png}
     \caption{{\bf Unit cell for Design B.} This design is still characterized by an alternation of open and closed cells.
@@
     \centering
-    \includegraphics[width=0.45\linewidth]{SFig4.png}
+    \includegraphics[width=0.4\linewidth]{SFig4.png}
     \caption{{\bf Unit cell for Design C.} This design consists of a square grid with all cells being reinforced by diagonal trusses passing through the nodes.
@@
     \centering
-    \includegraphics[width=0.45\linewidth]{SFig5.png}
+    \includegraphics[width=0.4\linewidth]{SFig5.png}
     \caption{{\bf Unit cell for Design D.} This design consists of a square grid without diagonal reinforcement.
@@
 
+\section{Numerical Model}
+The finite element analysis presented in this article was conducted using the {\it ABAQUS 2017} implicit scheme.
+For all buckling analysis, we performed a linear stability buckling analysis and for all stiffness results we performed a linear elastic analysis.
+All elements are assumed to be 1D Timoshenko beam elements and all beam crossings are assumed to be welded joints.
+The model assumes a linear elastic material with standard Poisson's ratio $\nu=0.3$ and a non-dimentionalized elastic Young's modulus of $E=1$ (resulting in all values of stresses being non-dimentionalized by the material's Young's modulus).
+
+For the boundary conditions, we treat the geometries from the previous section as Representative Volume Elements (RVE) which are used to generate infinite structures.
+To implement the infinite numerical scheme, we define periodic boundary conditions along the perimeter of the RVE.
+Each RVE is subjected to a macroscopic deformation gradient, so that
+\begin{equation}
+	\boldmath u_B-u_A=(\bar{F}-I)(X_B-X_A),
+\end{equation}
+where $A$ and $B$ are a pair of nodes located on the opposite side of the structure and $\boldmath u_X$ represents the finite displacement of node $X$.
+The deformation boundary conditions can thus be imposed on the deformation gradient given by $(\mathbold{\bar{F}-I})$, which are viewed as generalized degrees of freedom operationally applied using a set of virtual nodes.
+\citep{danielson2002}.
+
+For each instance, seeding of the mesh was chosen to be at least 1/10 of the minimum beam length.
+
 \section{Optimization Analysis}
-For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.
+For the optimization portion of the work presented in the main article we use a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.
 The algorithm parameters (continuous input variables) are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction.
 Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements.
-For simulations with odd number of diagonal re-enforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+
 The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and diagonal separation$=0.5*L$.
-The optimization is run under a uniaxial loading condition in the direction parallel non-diagonal elements.
-In order to maintain reasonable computational complexity, the optimization was run on the unit cell constructed using one dimensional beams with node seeding of at least 1/20 of each beam length.
 The covariance matrix is initialized uniformly with standard deviation ($\sigma$) half of the domain space, which are normalized to remain between 0 and 1.
+The optimization is run for uniaxial loading condition in the direction parallel to the non-diagonal elements.
 
-In optimization results presented in the main article, we seek to maximize the singly objective target function for the critical buckling load.
-However, an equivalent analysis is performed to maximize the critical buckling strain as the target response.
-This analysis is presented in \cref{BucklingOptimization}.
-Similarly, the same analysis was performed on the elastic stiffness of the structure and yields a trivial result leading to maximizing the allocation of mass to the non-diagonal elements through the mass ratio.
+For the optimization results presented in the main article, we seek to maximize the critical buckling load using a single objective target function.
+However, an equivalent analysis is performed to maximize the critical buckling strain as the target response (\cref{BucklingOptimization}.)
 
@@
 
+\section{Circular Cross-Section Results}
+
 \section{Parameter Exploration}
@@
 
-\section{Additional numerical results}
-\mf{Write this section}
-
- In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value.
-This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.
+\section{Local and Global Instabilities}
+For infinite structures modeled using period boundary conditions, it is important to realize the distinction between local instabilities (i.e.
+instabilities with wavelength that are of the order of the size of the unit cell) and global instabilities (i.e.
+instabilities with large wavelengths in comparison to the size of the unit cell).
+In order to capture any global instabilities potentially missed by using the minimum RVE, we further develop our period numerical studies to include RVEs that tessellate the unit cells presented in \cref{sec:designs}.
+For all designs, we increase the domain size by tessellating up to 20x20 unit cells (creating a base grid of 40x40 square cells).
+For Designs A-C, the critical bucking strain and modes do not change for larger unit cells suggesting that there does not exist a global instability.
+Whereas, for Design D, as we increase the domain size (add more unit cells to the tessellation), the instability  wavelength continues to grow with the size of the domain, suggesting a global macroscopic instability that depends on the size of the domain.
+The critical buckling strain is inversely proportional with the RVE size and approaches 0 in the limit that the RVE size approaches infinity.
+It is well known and documented that Design D is unstable when loaded in uniaxial compression.
 
 \mf{Write about minimum RVE -- how do I convey this message? Should I produce a plot or just verbally state this check?}
-
-\section{Analytical Derivation Of Uniaxial Stiffness}
-\mf{Write this section}
-%\subsection{Analytical Derivation of Critical Buckling Strain}
 

### 2018-09-24 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 \section{Circular Cross-Section Results}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\linewidth]{SFig7.pdf}
+	\caption{{\bf Circular Cross Section Results.} The color of the lines correspond to it's respective design color below the plots.
+(a) Shows the linear elastic stiffness of the different designs as a function of the loading angle.
+All structures except for the design without diagonal reinforcement have the same stiffness.
+(b) Shows the critical buckling strain for varying loading angle.
+For all angles, Design A outperforms other diagonally reinforced designs.
+These results show that the two diagonal benefit persists beyond a particular beam cross-section design.}
 
+	\label{CircularCrossSection}
+\end{figure}
 \section{Parameter Exploration}
@@
 \mf{Write this section}
+\subsection{Rectangular Cross Section}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.9\linewidth]{SFig8.pdf}
+	\caption{{\bf Circular Cross Section Results.} gsdfgs}
+	\label{CircularCrossSectionParameter}
+\end{figure}
+
 \subsection{Cylindrical Cross Section}
-\subsection{Rectangular Cross Section}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.9\linewidth]{SFig9.pdf}
+	\caption{{\bf Circular Cross Section Results.} gsdfgs}
+	\label{SquareCrossSectionParameter}
+\end{figure}
 

### 2018-09-24 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 \section{Circular Cross-Section Results}
+The results presented here complement that found on the main article and show that the structural benefit for the Design A persists when using a different cross-section for the structure.
+We show that for varying loading angles all of the diagonally reinforced designs provide the same stiffness, but Design A consistently provides the best resistance to buckling.
+
 \begin{figure}[H]
@@
 \section{Parameter Exploration}
-Include the graphs showing critical buckling strain as a function of parameters.\\
-\mf{Add figures for the diagonal}\\
-\mf{Add figure for $\lambda$ exploration}\\
-\mf{Make sure to include this analysis for both, rectangular cross-section and circular cross-section}\\
-\mf{Write this section}
+In order to survey the design space of the double diagonal construction we explore parametric simulations for 2 variables: diagonal separation and mass ratio.
+For each of these separate analysis, we maintain the sponge geometry as our base geometry and vary only the respective variable.
+
 \subsection{Rectangular Cross Section}
+This section shows the results when using a rectangular cross-section for the truss members.
+From \cref{SquareCrossSectionParameter}(a) it is apparent that there exists an optimum for the diagonal separation that occurs when the spacing between diagonals are approx 0.2 of the horizontal distance between vertical struts.
+This optimum value also persists when varying the loading angle.
+From \cref{SquareCrossSectionParameter}(b) it can be seen that the linear stiffness is symmetrically and almost purely dependent on the mass ratio allocated to diagonal versus non-diagonal elements.
+Comparing this figure to \cref{CircularCrossSectionParameter}(b) we can also see that even the design cross-section does not change the linear stiffness behavior.
+
 \begin{figure}[H]
@@
 	\includegraphics[width=0.9\linewidth]{SFig8.pdf}
-	\caption{{\bf Circular Cross Section Results.} gsdfgs}
-	\label{CircularCrossSectionParameter}
+	\caption{{\bf Rectangular Cross Section Parameter Exploration.} For each of these plots, we vary a single parameter while maintaining the base sponge inspired geometry constant.
+The gray line indicates the sponge design parameter.
+(a) Shows the critical buckling strain for varying spacing between diagonals.
+(b) Shows the structural stiffness of the geometry as we vary the mass ratio $\lambda$.
+(c) Shows the critical buckling strain of the geometry as we vary the mass ratio $\lambda$.}
+	\label{SquareCrossSectionParameter}
 \end{figure}
 
-\subsection{Cylindrical Cross Section}
+\subsection{Circular Cross Section}
 \begin{figure}[H]
@@
 	\includegraphics[width=0.9\linewidth]{SFig9.pdf}
-	\caption{{\bf Circular Cross Section Results.} gsdfgs}
-	\label{SquareCrossSectionParameter}
+	\caption{{\bf Circular Cross Section Parameter Exploration.} For each of these plots, we vary a single parameter while maintaining the base sponge inspired geometry constant.
+The gray line indicates the sponge design parameter.
+(a) Shows the critical buckling strain for varying spacing between diagonals.
+(b) Shows the structural stiffness of the geometry as we vary the mass ratio $\lambda$.
+(c) Shows the critical buckling strain of the geometry as we vary the mass ratio $\lambda$.}
+	\label{CircularCrossSectionParameter}
 \end{figure}

### 2018-09-24 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 
-\subsection{Rectangular Cross Section}
+\subsection{Rectangular Cross Section}\label{sec:rectparam}
 This section shows the results when using a rectangular cross-section for the truss members.
@@
 Comparing this figure to \cref{CircularCrossSectionParameter}(b) we can also see that even the design cross-section does not change the linear stiffness behavior.
+\cref{SquareCrossSectionParameter}(c) shows that there exists two optimum mass ratios, one where more material is allocated to the diagonal and one where there is more material allocated to non-diagonals.
+Note that if more matterialk
 
@@
 \subsection{Circular Cross Section}
+For this cross-section we stay consistent with the natural sponge, and we can see from \cref{CircularCrossSectionParameter} the overall behavior has second order differences for (a), no change for (b) and a large difference in relative magnitude for (c).
+Furthermore, we can see from the comparison between this section and \cref{sec:rectparam} that by using a rectangular cross-section we can tune the mass ratio to achieve an overall structure with higher critical buckling strength.
+
 \begin{figure}[H]

### 2018-09-28 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
Katia Edits

@@
 For the circular cross-section case  we stay consistent with the geometry found within the biological species.
-However, since slender members with such cross-section easily buckle out-of-plane when subject to compression loading we mitigate the issue by pursuing a plane-strain representative structure.
-To do so, we introduce the rectangular cross-section, where it's depth to slenderness ratio is sufficiently large to avoid out-of-plane deformation.
+However, since slender members with such circular cross-section easily buckle out-of-plane when subject to compression loading, we also consider  rectangular cross-sections and choose the depth to slenderness ratio  sufficiently large to avoid out-of-plane deformation.
 
-In the subsequent sections, we describe in detail the unit cells for four different designs (Designs A-D), and provide the derivations for each geometry cross-section values.
+In the subsequent sections, we describe in detail the unit cells for four different designs (Designs A-D), and provide the derivations for each geometry cross-section characteristics.
 
@@
 \subsubsection{Rectangular cross-section}
+\KB{clearly specify the dimension of the cross section. Don't use the diameter as dimension. It is confusing}
+\KB{I don't get the first sentence. We need to explain better}
 From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$.
-Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
+Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively: \KB{why do we have $D_{A,nd}$?
+is this a diameter?
+what is $t$?
+is it the in-plane or out-of-plane thickness?}
 \begin{equation}
@@
 \subsubsection{Circular cross section}
-It follows that the non-diagonal and diagonal volumes are given by
+For this design with circular cross section the non-diagonal and diagonal volumes are given by
 \begin{equation}
@@
 \subsubsection{Rectangular cross section}
-
+\KB{don't use $D$ as a dimension of the cross section}
 For this design the volume of non-diagonal elements remain the same, namely,
@@
 \subsubsection{Circular cross section}
-For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
+For this design with circular cross section, the non-diagonal and diagonal volumes of the unit cell are given by:
 \begin{equation}
@@
 \section{Numerical Model}
-The finite element analysis presented in this article was conducted using the {\it ABAQUS 2017} implicit scheme.
-For all buckling analysis, we performed a linear stability buckling analysis and for all stiffness results we performed a linear elastic analysis.
-All elements are assumed to be 1D Timoshenko beam elements and all beam crossings are assumed to be welded joints.
-The model assumes a linear elastic material with standard Poisson's ratio $\nu=0.3$ and a non-dimentionalized elastic Young's modulus of $E=1$ (resulting in all values of stresses being non-dimentionalized by the material's Young's modulus).
+The finite element analysis presented in this article were conducted using the {\it ABAQUS/Standard}.
+All models were constructed using 1D Timoshenko beam  (ABAQUS element type \KB{B??}) and all beam crossings were assumed to be welded joints.
+For each instance, seeding of the mesh was chosen to be at least 1/10 of the minimum beam length.
+The response of the material was captured using a linear elastic model with  Poisson's ratio $\nu=0.3$ and  Young's modulus of $E$.
 
-For the boundary conditions, we treat the geometries from the previous section as Representative Volume Elements (RVE) which are used to generate infinite structures.
-To implement the infinite numerical scheme, we define periodic boundary conditions along the perimeter of the RVE.
-Each RVE is subjected to a macroscopic deformation gradient, so that
-\begin{equation}
-	\boldmath u_B-u_A=(\bar{F}-I)(X_B-X_A),
+To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  a Representative Volume Element (RVE) \KB{add a schematic of the RVEs that you used}.
+To subject the RVE to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries  \citep{Danielsson2002JMPS, Bertoldi2008PRB}
+\begin{equation} \label{EQN:DefInfPeriodicFEM}
+    u_\alpha^{A_i} -u_\alpha^{B_i} = (\overline{F}_{\alpha\beta}-\delta_{\alpha\beta})(X_\beta^{A_i}-X_\beta^{B_i}), \quad i=1,\,2,\,....,\,N
 \end{equation}
-where $A$ and $B$ are a pair of nodes located on the opposite side of the structure and $\boldmath u_X$ represents the finite displacement of node $X$.
-The deformation boundary conditions can thus be imposed on the deformation gradient given by $(\mathbold{\bar{F}-I})$, which are viewed as generalized degrees of freedom operationally applied using a set of virtual nodes.
-\citep{danielson2002}.
+where $\delta_{\alpha\beta}$ is the Kronecker delta, $u_\alpha^{A_i}$  and $u_\alpha^{B_i}$ ($\alpha=1,\,2$) are displacements of points periodically located on the boundary of the unit cell.
+Moreover, $N$ denotes the number of pairs of nodes periodically located on the boundary of the unit cell.
+For the square lattice considered in this study $N=2$ when the unit cell is considered, while $N=4$ and $N=6$ for the enlarged $2\times2$ and $1\times5$ unit cells, respectively.
+Note that  the components of $\overline{\mathbf{F}}$ can be conveniently prescribed within the finite element framework using a set of virtual nodes.
+The corresponding  macroscopic first Piola-Kirchhoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{Danielsson2002JMPS, Bertoldi2008PRB}.
 
-For each instance, seeding of the mesh was chosen to be at least 1/10 of the minimum beam length.
+In this study the following macroscopic loading condition was considered: 
+\KB{add $\overline{\mathbf{F}}$ that you use in the analysis? this should be the most general case. How o you deal with the loading orientation $\theta$? }
+
+ To determine the initial stiffness of the structures  we performed a linear elastic analysis.
+\KB{ Which one to calculate the stiffness?
+Explain how do you conduct the analysis with different loading orientations}
+
+For all buckling analysis, we performed a linear stability buckling analysis \KB{did you use *buckling or *frequency? Did you account for the effect of pre-deformation}
+\KB{Since buckling may alter the periodicity of the
+structure, we considered super cells consisting of $m \times n$
+undeformed RVEs subjected to periodic boundary
+conditions and calculated the critical strain for each of
+them. The critical strain of the infinite periodic structure is
+then defined as the minimum critical strain on all considered
+super cells. \KB{it would be good to add this plot for each Design. Similar to Fig. 2c in "Complex Ordered Patterns in Mechanical Instability Induced Geometrically
+Frustrated Triangular Cellular Structures"}}
+
+\KB{how do you account for different $\theta$?}
+
+\KB{comment on the results}
+
+%For infinite structures modeled using period boundary conditions, it is important to realize the distinction between local instabilities (i.e.
+instabilities with wavelength that are of the order of the size of the unit cell) and global instabilities (i.e.
+instabilities with large wavelengths in comparison to the size of the unit cell).
+In order to capture any global instabilities potentially missed by using the minimum RVE, we further develop our period numerical studies to include RVEs that tessellate the unit cells presented in \cref{sec:designs}.
+For all designs, we increase the domain size by tessellating up to 20x20 unit cells (creating a base grid of 40x40 square cells).
+For Designs A-C, the critical bucking strain and modes do not change for larger unit cells suggesting that there does not exist a global instability.
+Whereas, for Design D, as we increase the domain size (add more unit cells to the tessellation), the instability  wavelength continues to grow with the size of the domain, suggesting a global macroscopic instability that depends on the size of the domain.
+The critical buckling strain is inversely proportional with the RVE size and approaches 0 in the limit that the RVE size approaches infinity.
+It is well known and documented that Design D is unstable when loaded in uniaxial compression.
+
+%\mf{Write about minimum RVE -- how do I convey this message? Should I produce a plot or just verbally state this check?}
 
 \section{Optimization Analysis}
-For the optimization portion of the work presented in the main article we use a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}.
-The algorithm parameters (continuous input variables) are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction.
+In an effort to identify lattice configurations resulting in large critical loads, we used  a Python
+implementation of the Covariance Matrix Adaptation Evolution
+Strategy (CMA-ES) \citep{hansen2003}.  CMA-ES is an evolutionary
+algorithm that is used to solve optimization/inverse problems
+by iteratively solving several forward problems to adjust
+a covariance matrix of the solution. CMA-ES is a derivative
+free algorithm, well suited for optimization problems of
+high dimensionality.
+
+\KB{this is not very clear. are we optimizing structures with circular or rectangular cross section? Which parameters do we keep fixed? Add a schematic}
+
+The
+parameters that need to be determined to achieve large critical loads are \KB{list all of them}
+\begin{itemize}
+\item the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and \item the separation of the even numbered diagonal elements from the non-diagonal junction.
 Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements.
 For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+\end{itemize}
 
@@
 However, an equivalent analysis is performed to maximize the critical buckling strain as the target response (\cref{BucklingOptimization}.)
+
+\KB{add expression for the target function}
 
@@
 \section{Local and Global Instabilities}
-For infinite structures modeled using period boundary conditions, it is important to realize the distinction between local instabilities (i.e.
-instabilities with wavelength that are of the order of the size of the unit cell) and global instabilities (i.e.
-instabilities with large wavelengths in comparison to the size of the unit cell).
-In order to capture any global instabilities potentially missed by using the minimum RVE, we further develop our period numerical studies to include RVEs that tessellate the unit cells presented in \cref{sec:designs}.
-For all designs, we increase the domain size by tessellating up to 20x20 unit cells (creating a base grid of 40x40 square cells).
-For Designs A-C, the critical bucking strain and modes do not change for larger unit cells suggesting that there does not exist a global instability.
-Whereas, for Design D, as we increase the domain size (add more unit cells to the tessellation), the instability  wavelength continues to grow with the size of the domain, suggesting a global macroscopic instability that depends on the size of the domain.
-The critical buckling strain is inversely proportional with the RVE size and approaches 0 in the limit that the RVE size approaches infinity.
-It is well known and documented that Design D is unstable when loaded in uniaxial compression.
-
-\mf{Write about minimum RVE -- how do I convey this message? Should I produce a plot or just verbally state this check?}
 

### 2018-09-29 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
added figures after katias suggestions and made progress on text

@@
 \subsubsection{Rectangular cross-section}
-\KB{clearly specify the dimension of the cross section. Don't use the diameter as dimension. It is confusing}
-\KB{I don't get the first sentence. We need to explain better}
-From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$.
-Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively: \KB{why do we have $D_{A,nd}$?
-is this a diameter?
-what is $t$?
-is it the in-plane or out-of-plane thickness?}
+
+Based on biological observations we can infer the ratio of non-diagonal to diagonal strut cross-sectional area to be  $A_{nd}/A_d\approx 1.4$.
+Using this information, we can develop the total volume ratio of non-diagonal to diagonal members to be $V_{nd}/V_d\approx 1.4$ assuming a uniform in-plane thickness $h$ for all struts.
+Using our sponge geometry interpretation we can obtain the following two expressions relating the members volume and thickness $T$:
 \begin{equation}
-	V_{A,nd}=8LD_{A,nd}t
+	V_{A,nd}=8LT_{A,nd}h
 \end{equation}
 \begin{equation}
-	V_{A,d}=8\sqrt{2}LD_{A,d}t
+	V_{A,d}=8\sqrt{2}LT_{A,d}h
 \end{equation}
-Let us enter this into our biological observation, which yields:
+where $V_{A,nd}$, $T_{A,nd}$,$V_{A,d}$ and $T_{A,d}$ are the non-diagonal volume, strut thickness and diagonal volume and strut thickness for Design A, respectively.
+Let us enter this relationship into our biological observation, which yields:
 \begin{equation}
-	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
+	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LT_{A,nd}h}{8\sqrt{2}LT_{A,d}h}
 \end{equation}
@@
 \begin{equation}
-	D_{A,nd}=2D_{A,d}
+	T_{A,nd}=2T_{A,d}
 \end{equation}
@@
 However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell.
+
 \subsubsection{Circular cross section}
@@
 \subsubsection{Rectangular cross section}
-\KB{don't use $D$ as a dimension of the cross section}
 For this design the volume of non-diagonal elements remain the same, namely,
 \begin{equation}
-		V_{B,nd}=8LD_{B,nd}t.
+		V_{B,nd}=8LT_{B,nd}h.
 \end{equation}
@@
 \begin{equation}
-		V_{B,d}=4\sqrt{2}LD_{A,d}t.
+		V_{B,d}=4\sqrt{2}LT_{A,d}h.
 \end{equation}
@@
 \begin{equation}
-		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
+		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LT_{B,nd}h}{4\sqrt{2}LT_{B,d}h}
 \end{equation}
@@
 \begin{equation}
-	D_{B,nd}=D_{B,d}
+	T_{B,nd}=T_{B,d}
 \end{equation}
@@
 \begin{equation}
-	D_{B,nd}=D_{A,nd}
+	T_{B,nd}=T_{A,nd}
 \end{equation}
@@
 \begin{equation}
-D_{A,nd}=2D_{A,d}
+T_{A,nd}=2T_{A,d}
 \end{equation}
@@
 \begin{equation}
-D_{C,nd}=D_{A,nd}
+T_{C,nd}=T_{A,nd}
 \end{equation}
@@
 \begin{equation}
-	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
+	V_{A,nd}+V_{A,d}=(T_{A,nd}+\sqrt{2}T_{A,d})8Lh
 \end{equation}
@@
 \begin{equation}
-V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}
+V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LhT_{A,nd}
 \end{equation}
@@
 \begin{equation}
-\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}=8LtD_{D,nd}
+\left(1+\frac{1}{\sqrt{2}}\right)8LhT_{A,nd}=8LhT_{D,nd}
 \end{equation}
@@
 \begin{equation}
-	D_{D,nd}=\left(1+\frac{1}{\sqrt{2}}\right)D_{A,nd}
+	T_{D,nd}=\left(1+\frac{1}{\sqrt{2}}\right)T_{A,nd}
 \end{equation}
@@
 The finite element analysis presented in this article were conducted using the {\it ABAQUS/Standard}.
-All models were constructed using 1D Timoshenko beam  (ABAQUS element type \KB{B??}) and all beam crossings were assumed to be welded joints.
+All models were constructed using 1D Timoshenko beam  (ABAQUS element type B22) and all beam crossings were assumed to be welded joints.
 For each instance, seeding of the mesh was chosen to be at least 1/10 of the minimum beam length.
@@
 
-To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  a Representative Volume Element (RVE) \KB{add a schematic of the RVEs that you used}.
-To subject the RVE to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries  \citep{Danielsson2002JMPS, Bertoldi2008PRB}
+To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  a Representative Volume Element (RVE) as seen in \cref{RVE}.
+To subject the RVE to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries  \citep{danielsson2002, bertoldi2008}
 \begin{equation} \label{EQN:DefInfPeriodicFEM}
@@
 Note that  the components of $\overline{\mathbf{F}}$ can be conveniently prescribed within the finite element framework using a set of virtual nodes.
-The corresponding  macroscopic first Piola-Kirchhoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{Danielsson2002JMPS, Bertoldi2008PRB}.
+The corresponding  macroscopic first Piola-Kirchhoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{danielsson2002, bertoldi2008}.
 
@@
 
+\begin{figure}[H]
+    \centering
+	\includegraphics[width=0.6\linewidth]{SFig10.png}
+	\caption{{\bf RVE used for the different designs.} This figure portrays the different RVEs used for the different designs considered.
+Each unit cell is shifted by half of a square cell to the horizontally and half of a square cell  vertically.
+Periodic boundary conditions are applied on the nodes that intersect with the light gray dashed line.}
+	\label{RVE}
+\end{figure}
+
 \section{Optimization Analysis}
@@
 
-The
-parameters that need to be determined to achieve large critical loads are \KB{list all of them}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.25\linewidth]{SFig11.png}
+	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals.
+For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$.
+Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.
+}
+	\label{OptimizationSchematic}
+\end{figure}
+
+The parameters that need to be determined to achieve large critical loads are: the number of diagonals, the mass ratio ($\lambda$) of non-diagonals to diagonal, and the separation between each even set of diagonals.
+For this optimization problem, the number of optimization variables increases with the number of diagonal incorporated in the model.
+Thus, the total number of parameters are $1+(N-1) + (N \text{ mod } 2)/N$ for a given optimization instance with $N$ number of diagonals.
+An illustration for an optimization structure with $N=5$ is shown \cref{OptimizationSchematic}.
+
 \begin{itemize}
@@
 % Bibliography
-\nocite{aizenberg2005}
-\nocite{deshpande2001}
-\nocite{miserez2008}
-\nocite{weaver2010}
+%\nocite{aizenberg2005}
+%\nocite{deshpande2001}
+%\nocite{miserez2008}
+%\nocite{weaver2010}
 

### 2018-09-30 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
added figures after katias suggestions and made progress on text
.

@@
 For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$.
-Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.
-}
+Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.}
 	\label{OptimizationSchematic}
@@
 \KB{add expression for the target function} 
+\\\mf{The target function is simply 1/ buckling reaction force, should we just keep this verbally instead of mathematically?}
 
@@
 \section{Local and Global Instabilities}
+\mf{How should I go about describing this here?}
 

### 2018-10-01 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 
-In this study the following macroscopic loading condition was considered: 
-\KB{add $\overline{\mathbf{F}}$ that you use in the analysis? this should be the most general case. How o you deal with the loading orientation $\theta$? }
+In this study the following uniaxial macroscopic loading condition was considered: \begin{equation}
+\overline{\mathbf{F}} = \begin{bmatrix}
+0&0\\
+0&-1
+\end{bmatrix} 
+\end{equation}
+
+In order to investigate the structure's buckling and elastic stiffness response in all directions, we rotate the unit cell model by angle $\theta$ and re-apply the above periodic boundary conditions using the rotated geometry coordinates.
+This approach allows us to rotate the base orthogonal coordinate system with respect to the unit cell.
 
 To determine the initial stiffness of the structures we performed a linear elastic analysis.
-\KB{ Which one to calculate the stiffness?
-Explain how do you conduct the analysis with different loading orientations}
+To calculate stiffness we use a displacement controlled simulation and extract the total reaction force from the model.
+To obtain the structures response to varying angle, we follow the simple procedure detailed in the previous paragraph to rotate the boundary condition.
+Knowing the displacement applied and the total reaction force, we can use this information to compute the structural stiffness.
 
@@
 \section{Optimization Analysis}
-In an effort to identify lattice configurations resulting in large critical loads, we used  a Python
-implementation of the Covariance Matrix Adaptation Evolution
-Strategy (CMA-ES) \citep{hansen2003}.  CMA-ES is an evolutionary
-algorithm that is used to solve optimization/inverse problems
-by iteratively solving several forward problems to adjust
-a covariance matrix of the solution. CMA-ES is a derivative
-free algorithm, well suited for optimization problems of
-high dimensionality.
-
-\KB{this is not very clear. are we optimizing structures with circular or rectangular cross section? Which parameters do we keep fixed? Add a schematic}
+In an effort to identify lattice configurations resulting in large critical loads, we used a Python implementation of the Covariance Matrix Adaptation Evolution Strategy (CMA-ES) \citep{hansen2003}.
+CMA-ES is an evolutionary
+algorithm that is used to solve optimization/inverse problems by iteratively solving several forward problems to adjust a covariance matrix of the solution.
+CMA-ES is a derivative free algorithm, well suited for optimization problems of high dimensionality and non-linear parameter topology.
 
@@
 
-The parameters that need to be determined to achieve large critical loads are: the number of diagonals, the mass ratio ($\lambda$) of non-diagonals to diagonal, and the separation between each even set of diagonals.
-For this optimization problem, the number of optimization variables increases with the number of diagonal incorporated in the model.
+The parameters that need to be determined to achieve large critical loads are: the number of diagonals, the mass ratio ($\lambda$) of non-diagonal to diagonal members, and the separation between each even set of diagonals.
+For this optimization problem, the number of optimization variables increases with the number of diagonals incorporated in the model.
 Thus, the total number of parameters are $1+(N-1) + (N \text{ mod } 2)/N$ for a given optimization instance with $N$ number of diagonals.
 An illustration for an optimization structure with $N=5$ is shown \cref{OptimizationSchematic}.
-
-\begin{itemize}
-\item the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and \item the separation of the even numbered diagonal elements from the non-diagonal junction.
-Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements.
-For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
-\end{itemize}
+Note, for simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
 

### 2018-10-01 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 
-For all buckling analysis, we performed a linear stability buckling analysis \KB{did you use *buckling or *frequency? Did you account for the effect of pre-deformation}
-\KB{Since buckling may alter the periodicity of the
-structure, we considered super cells consisting of $m \times n$
+For all buckling analysis, we performed a linear stability buckling analysis ({\tt *Buckling} command in ABAQUS input file).
+We perform this analysis for varying angles $\theta$ and extract the first mode critical buckling strain.
+Since buckling may alter the periodicity of the
+structure, we considered super cells consisting of $M \times M$
 undeformed RVEs subjected to periodic boundary
@@
 then defined as the minimum critical strain on all considered
-super cells. \KB{it would be good to add this plot for each Design. Similar to Fig. 2c in "Complex Ordered Patterns in Mechanical Instability Induced Geometrically
-Frustrated Triangular Cellular Structures"}}
+super cells.
 
-\KB{how do you account for different $\theta$?}
+\KB{it would be good to add this plot for each Design. Similar to Fig. 2c in "Complex Ordered Patterns in Mechanical Instability Induced Geometrically
+Frustrated Triangular Cellular Structures"} \mf{Yes, I like this plot and will work on creating it.}
 
-\KB{comment on the results}
+From the main article figure\#\# \mf{add reference to main article results figure} we conclude that for all $\theta$, Design A provides a higher overall buckling strength over the other compared designs, while not compromising it's overall stiffness in every direction.
 
@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\centering

### 2018-10-02 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
katia's new edits

@@
 \author[2]{James C. Weaver}
-\author[1,3,*]{Katia Bertoldi}
+\author[1,2,3]{Joanna Aizenberg}
+\author[1,2,3,*]{Katia Bertoldi}
 
@@
 
-\section{Designs Considered}\label{sec:designs}
-In this study, we focus on four different lattice configurations constrained to deform in an in-plane setting only.
-In an effort to conduct a fair performance comparison between the different designs, the total volume of material is conserved.
-Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry) as well as two different shapes for the cross section of all lattice members: circular and rectangular.
-For the circular cross-section case  we stay consistent with the geometry found within the biological species.
-However, since slender members with such circular cross-section easily buckle out-of-plane when subject to compression loading, we also consider  rectangular cross-sections and choose the depth to slenderness ratio  sufficiently large to avoid out-of-plane deformation.
+\section{Our four designs}\label{sec:designs}
+In this study, we focus on four different lattice configurations (Designs A, B, C and D) constrained to deform in an in-plane setting only.
+In an effort to conduct a fair performance comparison between the different designs, all four designs are characterized by the same total volume of material and a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry).
+Two different shapes are considered for teh cross section of the struts: circular and rectangular.
+For the circular cross-section case  we denote with $D_{\alpha,nd}$ and $D_{\alpha,d}$ the diameter of the non-diagonal (i.e.
+horizontal and vertical) and diagonal struts in the $\alpha$-th design, respectively, and neglect out-of-plane buckling.
+For the rectangular cross-sections we choose the depth $H$ and in-plane thickness $T_{\alpha,nd}$ and $T_{\alpha,d}$  to avoid out-of-plane deformation (i.e.
+we choose the  depth over thickness ratio sufficiently large).
+Finally, it is important to note that teh slenderness of the non-diagonal members in the $\alpha$-th design is chosen as
+\begin{equation}
+\frac{D_{\alpha,nd}}{L}=0.1,\;\;\;\text{and}\;\;\;\frac{T_{\alpha,nd}}{L}=0.1,
+\end{equation}
+for the case of circular and rectangular cross section, since this is the aspect ratio measured for the sponges (see \cref{sec:params}).
 
@@
 Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}).
+As for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell, so that they  intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes, where $L$ denote the length of the vertical/horizontal struts.
 \subsubsection{Circular cross section} 
-If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
+If we assume that cross section of all struts is circular, the projected-area and volume for the non-diagonal ($A_{A,nd}$ and $V_{A,nd}$) and diagonal ($A_{A,d}$ and $V_{A,d}$) members is given by
+\begin{equation}\label{A1}
+A_{A,nd}=8 L D_{A,nd},
+\end{equation}
 \begin{equation}\label{V1}
@@
 \end{equation}
-and
-\begin{equation}\label{A1}
-A_{A,nd}=8 L D_{A,nd},
-\end{equation}
-respectively. Note that in this study we use 
-\begin{equation}
-\frac{D_{A,nd}}{L}=0.1,
-\end{equation}
-since this is the aspect ratio measured for the sponges (see \cref{sec:params}).
-
-Moreover, as for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell.
-As such, the diagonals intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes and their  volume and projected area in the unit cell are
-\begin{equation}\label{V2}
-V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d},
-\end{equation}
-and
 \begin{equation}\label{A2}
@@
 \end{equation}
-respectively. Since the projected area ratio of the non-diagonal to diagonal elements in the sponge has been measured to be 
+and
+\begin{equation}\label{V2}
+V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d}.
+\end{equation}
+
+ Since the projected area ratio of the non-diagonal to diagonal elements in the sponge has been measured to be 
 \begin{equation}
@@
 
-Finally, it is important to note that in this study we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs to be equal to that of Design A, namely,
+Finally, it is important to note that in this study we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with circular cross section to be equal to that of Design A, namely,
 \begin{equation}\label{con1}
@@
 
-Based on biological observations we can infer the ratio of non-diagonal to diagonal strut cross-sectional area to be  $A_{nd}/A_d\approx 1.4$.
-Using this information, we can develop the total volume ratio of non-diagonal to diagonal members to be $V_{nd}/V_d\approx 1.4$ assuming a uniform in-plane thickness $h$ for all struts.
-Using our sponge geometry interpretation we can obtain the following two expressions relating the members volume and thickness $T$:
+If we assume that the cross section of all struts is rectangular, the projected-area  for the non-diagonal ($A_{A,nd}$) and diagonal ($A_{A,d}$) members is given by
 \begin{equation}
-	V_{A,nd}=8LT_{A,nd}h
+	A_{A,nd}=8LT_{A,nd}
 \end{equation}
+and
 \begin{equation}
-	V_{A,d}=8\sqrt{2}LT_{A,d}h
+	A_{A,d}=8\sqrt{2}LT_{A,d}
 \end{equation}
-where $V_{A,nd}$, $T_{A,nd}$,$V_{A,d}$ and $T_{A,d}$ are the non-diagonal volume, strut thickness and diagonal volume and strut thickness for Design A, respectively.
-Let us enter this relationship into our biological observation, which yields:
+where  $T_{A,nd}$ and $T_{A,d}$ are the non-diagonal and diagonal in-plane strut thickness for Design A, respectively.
+Since for the sponge   $A_{nd}/A_d\approx 1.4$, it follows that
 \begin{equation}
-	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LT_{A,nd}h}{8\sqrt{2}LT_{A,d}h}
+	T_{A,nd}=2T_{A,d}.
 \end{equation}
-Therefore, we can simplify this relationship as
-\begin{equation}
-	T_{A,nd}=2T_{A,d}
+
+Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely, \KB{update eqn}
+\begin{equation}\label{con1R}
+V_{\alpha,d}+V_{\alpha,nd}={V}_{A,T}= ??,
 \end{equation}
+with $\alpha=$ B, C and D.
+Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A \KB{update eqn}
+\begin{equation}\label{con2R}
+\frac{V_{\alpha,nd}}{V_{\alpha,d}}=???,
+\end{equation}
+with $\alpha=$ B and C.
 
@@
 \subsubsection{Rectangular cross section}
-For this design the volume of non-diagonal elements remain the same, namely,
+For this design with circular cross section the volume of the non-diagonal and diagonal mmebers  are given by
 \begin{equation}
-		V_{B,nd}=8LT_{B,nd}h.
+		V_{B,nd}=8LT_{B,nd}H.
 \end{equation}
-However, the volume for the diagonal elements will is different as a result the change in total length, namely,
+and
 \begin{equation}
-		V_{B,d}=4\sqrt{2}LT_{A,d}h.
+		V_{B,d}=4\sqrt{2}LT_{A,d}H.
 \end{equation}
-In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
-\begin{equation}
-		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LT_{B,nd}h}{4\sqrt{2}LT_{B,d}h}
-\end{equation}
-which simplifies to
+Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
 \begin{equation}
@@
 \subsubsection{Circular cross section}
-For this design with circular cross section, the non-diagonal and diagonal volumes of the unit cell are given by:
+For this design with circular cross section, the volume of the non-diagonal and diagonal members  of the unit cell are given by:
 \begin{equation}
@@
 \subsubsection{Rectangular cross section}
-Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
+For this design with circular cross section, the volume of the non-diagonal and diagonal members  of the unit cell are given by
+\KB{add eqns.}
+Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
 \begin{equation}
-T_{A,nd}=2T_{A,d}
+T_{A,nd}=2T_{A,d},
 \end{equation}
@@
 \begin{equation}
-T_{C,nd}=T_{A,nd}
+T_{C,nd}=T_{A,nd}.
 \end{equation}
@@
 \subsection{Design D} 
-Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}.
+Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}).
 As such, for this design we allocate the total material volume to the non-diagonal elements.
+Note that this design is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}.
 
@@
 \subsubsection{Rectangular cross-section}
-For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements.
-Thus we can formulate the total volume of Design A as
+Since
 \begin{equation}
-	V_{A,nd}+V_{A,d}=(T_{A,nd}+\sqrt{2}T_{A,d})8Lh
+V_{D,T}=V_{D,nd}=8 L T_{D,nd} H,
 \end{equation}
-Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
-\begin{equation}
-V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LhT_{A,nd}
-\end{equation}
-
-We can build a relationship between Design A and Design D through the total volume, such that the total volume of Design A is the same as the total volume of Design D
-\begin{equation}
-\left(1+\frac{1}{\sqrt{2}}\right)8LhT_{A,nd}=8LhT_{D,nd}
-\end{equation}
-which yields a relationship between the thicknesses of the non diagonal elements as
+using the constraint provided by \cref{con1R} we obtain
 \begin{equation}
@@
 
-To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  a Representative Volume Element (RVE) as seen in \cref{RVE}.
-To subject the RVE to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries  \citep{danielsson2002, bertoldi2008}
+To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  the unit cells shown  in \cref{RVE}.
+To subject the unit cells to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries by enforcing  \citep{danielsson2002, bertoldi2008}
 \begin{equation} \label{EQN:DefInfPeriodicFEM}
@@
 Moreover, $N$ denotes the number of pairs of nodes periodically located on the boundary of the unit cell.
-For the square lattice considered in this study $N=2$ when the unit cell is considered, while $N=4$ and $N=6$ for the enlarged $2\times2$ and $1\times5$ unit cells, respectively.
+\KB{what does the following sentence mean?
+Are you referring to the unit cells shown in Fig.
+6?}For the square lattice considered in this study $N=2$ when the unit cell is considered, while $N=4$ and $N=6$ for the enlarged $2\times2$ and $1\times5$ unit cells, respectively.
 Note that  the components of $\overline{\mathbf{F}}$ can be conveniently prescribed within the finite element framework using a set of virtual nodes.
-The corresponding  macroscopic first Piola-Kirchhoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{danielsson2002, bertoldi2008}.
+The corresponding  macroscopic first Piola-Kirchoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{danielsson2002, bertoldi2008}.
 
-In this study the following uniaxial macroscopic loading condition was considered: \begin{equation}
+To subject the structures to uniaxial compression, in this study the following uniaxial macroscopic loading condition was considered:\KB{are you sure?
+I believe we left $\lambda_x$ unset.
+Right?} \begin{equation}
 \overline{\mathbf{F}} = \begin{bmatrix}
 0&0\\
-0&-1
+0&\lambda_y
 \end{bmatrix} 
@@
 
-The parameters that need to be determined to achieve large critical loads are: the number of diagonals, the mass ratio ($\lambda$) of non-diagonal to diagonal members, and the separation between each even set of diagonals.
-For this optimization problem, the number of optimization variables increases with the number of diagonals incorporated in the model.
-Thus, the total number of parameters are $1+(N-1) + (N \text{ mod } 2)/N$ for a given optimization instance with $N$ number of diagonals.
-An illustration for an optimization structure with $N=5$ is shown \cref{OptimizationSchematic}.
-Note, for simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
+In this study we used CMA-ES to identify \begin{itemize} 
+\item the number of diagonals, $N$
+\item the volume ratio  of non-diagonal to diagonal members, $\lambda=V_{nd}/V_{d}$.
+\item  the separation between each even set of diagonals, $S_i$ (see Fig. \ref{OptimizationSchematic})
+\end{itemize}
+resulting in a lattice structure with the large critical load.
+For such optimization problem, the number of optimization variables increases with the number of diagonals incorporated in the model(i.e.
+the total number of parameters are $1+ \frac{1}{2}(N - (N \text{ mod } 2))$ for a given optimization instance with $N$ number of diagonals).
+Note, for simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry (see Fig.
+\ref{OptimizationSchematic}).
 
-The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and diagonal separation$=0.5*L$.
+The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and \KB{is thsi $S_i$??}diagonal separation$=0.5*L$.
 The covariance matrix is initialized uniformly with standard deviation ($\sigma$) half of the domain space, which are normalized to remain between 0 and 1.
-The optimization is run for uniaxial loading condition in the direction parallel to the non-diagonal elements.
+The optimization is run for uniaxial loading condition in the direction parallel to the vertical elements.
 
@@
 However, an equivalent analysis is performed to maximize the critical buckling strain as the target response (\cref{BucklingOptimization}.)
-
-\KB{add expression for the target function} 
-\\\mf{The target function is simply 1/ buckling reaction force, should we just keep this verbally instead of mathematically?}
 

### 2018-10-03 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.
katia's new edits

@@
-\documentclass[9pt,twocolumn,twoside]{fernandes_paper} 
+\documentclass[10pt,twocolumn,twoside]{fernandes_paper} 
 %Default opts:9pt,twocolumn,twoside
@@
 % \doublespace
-\linenumbers
-
-% \section{Introduction}
-% \subsection{Sponge Biology}
-% \subsection{History of Diagonally Reinforced Structures and Previous Works}
-% \subsection{Examples of Current Diagonally Reinforced Designs}
-
-% \section{Methods}
-% \subsection{Numerical Model Setup}
-% \subsection{Experimental Design}
-
-% \section{Results}
-% \subsection{Linear Stiffness for different Designs}
-%  \subsection{Buckling Resistance for Different Designs}
-% \subsection{Experimental Results}
-
-% \section{Discussion}
-% \subsection{Parameter Exploration}
-% \subsection{Parameter Optimization}
-
-% \section{Conclusion}
-
-% \section*{Acknowledgments}
-% \section*{References}
+%\linenumbers
 
@@
 	\includegraphics[width=0.9\textwidth]{Fig2}
-	\caption{text}\label{Fig2}
+	\caption{\textbf{Experimental Results.} (a) shows the different designs (Design A-D) considered in this analysis.
+(b) shows the undeformed experimental setup for the different designs.
+(c) shoes the post-buckling deformed state for each of the designs.
+(d) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.}\label{Fig2}
 \end{figure*}
@@
 	\end{center}
-	\caption{\textbf{text} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs.
+	\caption{\textbf{Structure Mechanical Response} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs.
 The designs considered are depicted below as Design A-D.} \label{Fig3}
@@
 
-% \section{Introduction}
-
-In modern day structural engineering, we find many
+\begin{figure*}[ht]
+	\captionsetup{width=0.8\textwidth}
+	\begin{center}
+		\includegraphics[width=0.7\textwidth]{Fig4}
+	\end{center}
+	\caption{\textbf{Critical Buckling Load Optimization Results.} (a) shows the optimal value of critical buckling load for varying number of diagonals.
+The color of each point represents the optimal mass ratio $\lambda$  parameter for that configuration.
+Note, we do not explicitly show the other optimal values for the diagonal spacing parameters.
+(b) shows the resulting deformed geometries for designs including one to four diagonals.} \label{Fig4}
+\end{figure*}
 
@@
 
-Hexactinellids, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m.
+\textit{Hexactinellids}, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m.
 Beyond their fracture propagation inhibiting material composition \citep{weaver2007}, they are perceived to exhibit large structural rigidity and strength against buckling.
-% Since they are primarily made of 'brittle silica', buckling strength may be a crucial property in making them resistant to impact and environmentally applied stresses. 
+
 Structurally, they are composed of a base square-grid architecture and regular ordering of vertical and horizontal struts that form the skeletal system.
@@
 This diagonal reinforcement design is conjectured to give the sponge greater buckling resistance and strength to localized damage then it would experience having a single diagonal reinforcement strut (with equivalent allocation of mass between the diagonal reinforcements.)
-% Analogous to the sponge, many engineering structures, such as buildings and bridges, exhibit diagonal reinforcement struts as a stability mechanism.
-Based on this similarity, we explore the following research question:
-The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?} \mf{this paragraph needs more citations and more avid lit.
-survey}
 
-In this paper, we present a numerical analysis of the structure deformation under various loading conditions as well as survey different arrangements within similar design space of the sponge via parametric studies.
-Through the various design iterations we look for the critical buckling strain and the elastic load carrying capacity.
-Finally, we compare the results obtained from the sponge design to what is typically used in engineering structures and validate our results with experimental data.
+The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?}
 
-% involving the development of truss structures with efficient diagonal reinforcement, is inspired by glass sponges (primarily made of silica) commonly known as the Venus Flower Basket.
-These deep sea sponges exhibit a regular and periodic truss structure which biologists for many years have stipulated to be an optimized design for structural and hydrodynamic purposes.
-To investigate these claims, I have built numerical and analytical models analyzing the buckling effects of the sponge designs by tuning various geometrical parameters and analyzing its effect on the buckling response.
-Thus far, we have found the the double diagonal reinforced sponge design delays the onset of structural buckling by a factor of twice the applied strain in comparison to the traditional designs.
-We believe that this intricate approach inspired by the sponge can improve the way we construct buildings and bridges, which will give the opportunity to build taller buildings and bridges that use less material and are stronger.
+In this paper, we present a numerical and experimental mechanical analysis of multiple  structures under various loading conditions as well as survey different arrangements within similar design space via parametric studies.
+Furthermore, we incorporate an optimization algorithm on our parameter space to obtain an optimal design to withstand (via critical buckling load) uniaxial loading parallel to the non-diagonal truss members.
+For this analysis, we restrict ourselves to four designs considered (Design A-D) that are shown along the top of \cref{Fig2}.
+As part of our analysis, we study the critical buckling strain and the elastic load carrying capacity of these structures.
 
-% \section{Results}
-% \subsection{Linear Stiffness for different Designs}
-In this section, we compare the linear stiffness for the different arrangements using Finite Element Analysis (FEA).
-For each design, we consider a Representative Volume Element (RVE) and apply periodic boundary conditions along the edges of the unit cell domains depicted along the bottom of \cref{Fig3}.
-In our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections.
+Numerically, we compare the linear stiffness and critical buckling strain for each of the different designs using Finite Element Analysis (FEA).
+To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI).
+For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections.
 In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.
@@
 
-% \subsection{Buckling Resistance for Different Designs}
 The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements.
@@
 
-% \subsection{Experimental Validation}
 For the experimental validation, we conducted various uniaxial compression tests of 3D printed planar beam structures (with circular cross-section) on an Instron 5566.
@@
 The experimental set-up consisted of two \hl{1 inch} thickness acrylic sheets bolted together along the perimeter and spaced from each other by the thickness of the 3D printed specimen.\\\\\hl{[...]} \mf{more to go here when we have experimental setup and data}
-
-% \figll{Buckling_Separation}{0.9}{Critical buckling strain for varying diagonal separation in Design A parametric sweep.
-The normalized separation distance $C/L$ indicates how far apart the diagonals are from each other.
-For $C/L=0$ we obtain the Design B and for $C/L=0.5$ we obtain the diagonals crossing half way through the length of the non-diagonal trusses ($L$).
-For all values in this sweep, we assume a mass ratio equivalent to that of the sponge.
-The dashed gray line indicates the assumed diagonal separation of the sponge of $C/L=1/(\sqrt{2}+2)$.}
 
@@
 
-% \section{Mass Ratio Parametrization}
-% \figll{Stiffness_MassRatio}{0.9}{Structural stiffness for varying mass ratio parametric sweep.
-Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements.
-For all values in this graph, the design is kept constant while the cross-sectional area of each element is scaled accordingly.
-For all values of $\lambda$, the total mass of all elements are kept constant.
-The different colored lines correspond to different loading angles of the unit cell.
-The gray dashed line indicates the measured mass ratio for the sponge.
-For all values of this sweep, we assume Design A geometry although the linear stiffness does not depend on the design.}
-
-% \figll{Buckling_MassRatio}{0.9}{Critical buckling strain for varying mass ratio parametric sweep.
-Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements.
-For all values in this graph, the design is kept constant while the cross-sectional area of each element is scaled accordingly.
-For all values of $\lambda$, the total mass of all elements are kept constant.
-The different colored lines correspond to different loading angles of the unit cell.
-The gray dashed line indicates the measured mass ratio for the sponge.
-For all values of this sweep, we assume Design A geometry.}
-
-% \section{Diagonal Separation Parametrization}
-
 To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements.
@@
 
-{Fig5}(b) shows the structural stiffness as a function of $\lambda$ for various loading angles (corresponding to the respective colors).
-In this figure, we see that for small $\lambda$ the structure is stiffer at $0^\circ$ loading and for large $\lambda$ the structure is stiffer at $45^\circ$.
-This is expected as we are allocating more material to its respective elements.
-Furthermore, we can see that at $\lambda=1$ we attain a point in the mass ratio where the stiffness is equal for all angles.
-That is because there is the same amount of material allocated to the diagonals as there is to the non-diagonal elements.
+%\section*{Acknowledgments}
 
-{Fig5}(c) illustrates the critical buckling strain as a function of $\lambda$.
-Here, we can clearly see that buckling strain is dependent on the mass ratio and behaves differently for all angles \hl{[...]} \mf{I am not sure what to say here (interpretation) about this graph.}
-
-% \section{Conclusion}
-% \hl{add writing}
-
-\section*{Funding Information}
-\hl{add writing}
 % No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship
-
-\section*{Acknowledgments}
-\hl{add writing} 
 % Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft
 
-%\begin{figure*}
-%	\captionsetup{width=0.8\textwidth}
-%	\begin{center}
-%		\includegraphics[width=0.7\textwidth]{Fig5}
-%	\end{center}
-%	\caption{\textbf{Parametric sweep results for different analysis.} (a) Critical buckling strain for varying diagonal separation in Design A.
-The normalized separation distance $C/L$ indicates how far apart the diagonals are from each other.
-For all values we assume a mass ratio to be equivalent to that of the sponge.
-The dashed gray line indicates the assumed diagonal separation of the sponge, namely $C/L=1/(\sqrt{2}+2)$.
-(b) Structural linear stiffness for varying mass ratio.
-Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements based on the volume element.
-For all values of $\lambda$, the total mass of all elements are kept constant.
-The gray dashed line indicates the measured mass ratio for the sponge.
-(c) Critical buckling strain for varying mass ratio.
-For all values of $\lambda$, the total mass of all elements are kept constant.
-The different colored lines correspond to different loading angles of the unit cell.
-The gray dashed line indicates the measured mass ratio for the sponge.
-For all values in (b) and (c), we assume Design A geometry.} \label{Fig5}
-%\end{figure*}
-
- Bibliography
 \nocite{*}

### 2018-10-03 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 In an effort to conduct a fair performance comparison between the different designs, all four designs are characterized by the same total volume of material and a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry).
-Two different shapes are considered for teh cross section of the struts: circular and rectangular.
-For the circular cross-section case  we denote with $D_{\alpha,nd}$ and $D_{\alpha,d}$ the diameter of the non-diagonal (i.e.
+Two different shapes are considered for the cross-section of the struts: circular and rectangular.
+For the circular cross-section case  we denote the diameters with $D_{\alpha,nd}$ and $D_{\alpha,d}$ of the non-diagonal (i.e.
 horizontal and vertical) and diagonal struts in the $\alpha$-th design, respectively, and neglect out-of-plane buckling.
@@
 we choose the  depth over thickness ratio sufficiently large).
-Finally, it is important to note that teh slenderness of the non-diagonal members in the $\alpha$-th design is chosen as
+Finally, it is important to note that the slenderness of the non-diagonal members in the $\alpha$-th design $\in [A,B,C]$ is chosen as
 \begin{equation}
-\frac{D_{\alpha,nd}}{L}=0.1,\;\;\;\text{and}\;\;\;\frac{T_{\alpha,nd}}{L}=0.1,
+\frac{D_{\alpha,nd}}{L}=0.1,\quad\text{and}\quad\frac{T_{\alpha,nd}}{L}=0.1,
 \end{equation}
@@
 
-In the subsequent sections, we describe in detail the unit cells for four different designs (Designs A-D), and provide the derivations for each geometry cross-section characteristics.
+In the subsequent sections, we describe in detail the unit cells for four different designs, and provide the derivations for each geometry cross-section characteristics.
 
@@
 \subsubsection{Circular cross section} 
-If we assume that cross section of all struts is circular, the projected-area and volume for the non-diagonal ($A_{A,nd}$ and $V_{A,nd}$) and diagonal ($A_{A,d}$ and $V_{A,d}$) members is given by
+If we assume that cross section of all struts is circular, the projected area and volume for the non-diagonal ($A_{A,nd}$ and $V_{A,nd}$) and diagonal ($A_{A,d}$ and $V_{A,d}$) members is given by
 \begin{equation}\label{A1}
@@
 
-Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely, \KB{update eqn}
+Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely,
 \begin{equation}\label{con1R}
-V_{\alpha,d}+V_{\alpha,nd}={V}_{A,T}= ??,
+{V}_{A,T}=V_{\alpha,d}+V_{\alpha,nd}= 8LH(T_{A,nd}+\sqrt{2}T_{A,d})=8LHT_{A,nd}\left(1+\frac{1}{\sqrt{2}}\right),
 \end{equation}
 with $\alpha=$ B, C and D.
-Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A \KB{update eqn}
+Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A,
 \begin{equation}\label{con2R}
-\frac{V_{\alpha,nd}}{V_{\alpha,d}}=???,
+\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\sqrt{2},
 \end{equation}
-with $\alpha=$ B and C.
+with $\alpha\in$ B and C.
 
@@
 \subsubsection{Rectangular cross section}
-For this design with circular cross section the volume of the non-diagonal and diagonal mmebers  are given by
+For this design with circular cross section the volume of the non-diagonal and diagonal members  are given by
 \begin{equation}
@@
 \begin{equation}
-		V_{B,d}=4\sqrt{2}LT_{A,d}H.
+		V_{B,d}=4\sqrt{2}LT_{B,d}H.
 \end{equation}
@@
 For this design with circular cross section, the volume of the non-diagonal and diagonal members  of the unit cell are given by
-\KB{add eqns.}
+\begin{equation}
+V_{C,nd}=8LT_{C,nd}H
+\end{equation}
+and
+\begin{equation}
+V_{C,d}=8\sqrt{2}LT_{C,d}H
+\end{equation}
 Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
 \begin{equation}
-T_{A,nd}=2T_{A,d},
+T_{C,nd}=2T_{C,d},
 \end{equation}
@@
 \begin{equation} \label{EQN:DefInfPeriodicFEM}
-    u_\alpha^{A_i} -u_\alpha^{B_i} = (\overline{F}_{\alpha\beta}-\delta_{\alpha\beta})(X_\beta^{A_i}-X_\beta^{B_i}), \quad i=1,\,2,\,....,\,N
+    \mathbf{u}_\alpha^{A_i} -\mathbf{u}_\alpha^{B_i} = (\mathbf{\overline{F}}_{\alpha\beta}-\delta_{\alpha\beta})(\mathbf{X}_\beta^{A_i}-\mathbf{X}_\beta^{B_i}), \quad i=1,\,2,\,....,\,K
 \end{equation}
-where $\delta_{\alpha\beta}$ is the Kronecker delta, $u_\alpha^{A_i}$  and $u_\alpha^{B_i}$ ($\alpha=1,\,2$) are displacements of points periodically located on the boundary of the unit cell.
-Moreover, $N$ denotes the number of pairs of nodes periodically located on the boundary of the unit cell.
-\KB{what does the following sentence mean?
-Are you referring to the unit cells shown in Fig.
-6?}For the square lattice considered in this study $N=2$ when the unit cell is considered, while $N=4$ and $N=6$ for the enlarged $2\times2$ and $1\times5$ unit cells, respectively.
+where $\delta_{\alpha\beta}$ is the Kronecker delta, $\mathbf{u}_\alpha^{A_i}$  and $\mathbf{u}_\alpha^{B_i}$ ($\alpha=1,\,2$) are displacements of points periodically located on the boundary of the unit cell.
+Moreover, $K$ denotes the number of pairs of nodes periodically located on the boundary of the unit cell.
 Note that  the components of $\overline{\mathbf{F}}$ can be conveniently prescribed within the finite element framework using a set of virtual nodes.
@@
 
-To subject the structures to uniaxial compression, in this study the following uniaxial macroscopic loading condition was considered:\KB{are you sure?
-I believe we left $\lambda_x$ unset.
-Right?} \begin{equation}
+To subject the structures to uniaxial compression, in this study the following uniaxial macroscopic loading condition was considered: \begin{equation}
 \overline{\mathbf{F}} = \begin{bmatrix}
-0&0\\
+\text{UNSET}&0\\
 0&\lambda_y
@@
 	\centering
-	\includegraphics[width=0.25\linewidth]{SFig11.png}
+	\includegraphics[width=0.2\linewidth]{SFig11.png}
 	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals.
@@
 
-The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and \KB{is thsi $S_i$??}diagonal separation$=0.5*L$.
+The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and diagonal separation for the even set of diagonals $S_i=0.5*L$.
 The covariance matrix is initialized uniformly with standard deviation ($\sigma$) half of the domain space, which are normalized to remain between 0 and 1.
@@
 \cref{SquareCrossSectionParameter}(c) shows that there exists two optimum mass ratios, one where more material is allocated to the diagonal and one where there is more material allocated to non-diagonals.
-Note that if more matterialk
 
@@
 \mf{How should I go about describing this here?}
-
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{SFig12.pdf}
+	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals.
+For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$.
+Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.}
+	\label{DomainSizeSchematic}
+\end{figure}
 % Bibliography

### 2018-10-03 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 then defined as the minimum critical strain on all considered
-super cells.
-
-\KB{it would be good to add this plot for each Design. Similar to Fig. 2c in "Complex Ordered Patterns in Mechanical Instability Induced Geometrically
-Frustrated Triangular Cellular Structures"} \mf{Yes, I like this plot and will work on creating it.}
+super cells. See \cref{DomainSize} for this detailed analysis.
 
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{SFig6.png}
+    \includegraphics[width=0.9\linewidth]{SFig6.pdf}
     \caption{{\bf Buckling Optimization.} Graph shows optimal value of critical buckling strain for varying number of diagonals.
@@
 	\includegraphics[width=0.7\linewidth]{SFig12.pdf}
-	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals.
-For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$.
-Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.}
-	\label{DomainSizeSchematic}
+	\caption{{\bf Global vs Local Instabilities.} This figure shows four identical analysis performed on the Designs A-D.
+For each analysis we obtain the critical buckling strain for varying $M\times M$ tessellations (y-axis) of the unit cell also for varying loading angle $\theta$ (x-axis).
+The color of a pixel corresponds to a value of critical buckling strain for that simulation.
+For each of the simulations, periodic boundary conditions are applied along the outer perimeter of the $M\times M$ structure.
+This plot conveys that for Designs A-C the prominent buckling mode is the local mode, whereas for Design D, the prominent mode is a global mode.
+Choosing a sufficiently large $M$ allows Design D to converge to a finite value for each $\theta$.}
+	\label{DomainSize}
 \end{figure}

### 2018-10-03 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.
Updates from ShareLaTeX

--- 2018-08-23 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
@@
 \newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}
+\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
 
@@
 
+\begin{figure*}[ht]
+	\captionsetup{width=0.8\textwidth}
+	\begin{center}
+		\includegraphics[width=0.7\textwidth]{Fig4}
+	\end{center}
+	\caption{\textbf{text} Figure showing (a) (b)} \label{Fig4}
+\end{figure*}
+
 % \section{Introduction}
 
-In modern day structural engineering, we find many
+\KB{Lattice structures with well-defined periodicity are ubiquitous
+ in synthetic structures and
+devices as they offer novel and unique
+properties, including light weight REFs,  high energy absorption REFs,
+ and the ability to control the propagation of both
+electromagnetic REFs and elastic waves REFs and heat flow REFs.} \KB{INTRODUCE THE SQUARE LATTICE AN SAY THAT IT IS VERY POPULAR BUT UNSTABLE. NEED FOR REINFORCEMENT}
 
-The design of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
+The design of diagonal reinorcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
 In modern day structures, we see many different iterations of new truss designs with various applications in mind but with what is the same fundamentally simple diagonal reinforcement technique.
@@
 
- Bibliography
 \nocite{*}

### 2018-10-03 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
overleaf

--- 2018-10-03 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
@@
 
-The design of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
+\KB{Lattice structures with well-defined periodicity are ubiquitous
+ in synthetic structures and
+devices as they offer novel and unique
+properties, including light weight REFs,  high energy absorption REFs,
+ and the ability to control the propagation of both
+electromagnetic REFs and elastic waves REFs and heat flow REFs.} \KB{INTRODUCE THE SQUARE LATTICE AN SAY THAT IT IS VERY POPULAR BUT UNSTABLE. NEED FOR REINFORCEMENT}
+
+The design of diagonal reinorcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
 In modern day structures, we see many different iterations of new truss designs with various applications in mind but with what is the same fundamentally simple diagonal reinforcement technique.
@@
 %\section*{Acknowledgments}
+% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship
+% James R. Rice. Mark Scott. Joost Vlassak. Chris Rycroft for insightful discussions
 
-% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship
-% Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft
-
-\nocite{*}
+%\nocite{*}
 \bibliography{refs}

### 2018-10-24 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
new version

--- 2018-10-03 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
@@
-\documentclass[10pt,twocolumn,twoside]{fernandes_paper} 
+\documentclass[9pt,twocolumn,twoside]{fernandes_paper} 
 %Default opts:9pt,twocolumn,twoside
@@
 
-\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}
+\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
+% \newcommand{\mf}[1]{\noindent\color{color2}\sethlcolor{cyan}\hl{$\Longrightarrow$#1}\normalcolor}
+\newcommand{\mf}[1]{\noindent\color{orange}{$\Longrightarrow$#1}\normalcolor}
 
-\title{Utilizing Design Principles from Glass Sponges for Structurally Robust Lattices}
+\title{Design Principles from Glass Sponges \\for Structurally Robust Lattices}
 
@@
 
-\keywords{Architected materials, truss structure, buckling resistance, glass sponges, bio-inspired engineering.}
-
-\usepackage{comment}
+\keywords{Architected materials, truss structure, square lattice, buckling resistance, glass sponges, bio-inspired engineering.}
 
@@
 % \doublespace
-%\linenumbers
+\linenumbers
 
-\begin{figure*}[ht]
+\begin{figure*}[!htb]
+    \centering
     \captionsetup{width=0.8\textwidth}
-\begin{center}
-\includegraphics[width=0.9\textwidth]{Fig1}
-\end{center}
+    \includegraphics[width=0.6\textwidth]{Fig1}
     \caption{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge.
 (c) Close up microscope image of the sponge.
-(d) Comparison between the idealized model (green and blue lines) and the sponge structure.
-(e) Representative volume element unit cell of the idealized model.}\label{Fig1}
+(d) Comparison between the idealized model (green and blue lines) and the sponge structure.}\label{Fig1}
 \end{figure*}
 
-\begin{figure*}[ht]
+\begin{figure*}[!htb]
 	\centering
@@
 	\includegraphics[width=0.9\textwidth]{Fig2}
-	\caption{\textbf{Experimental Results.} (a) shows the different designs (Design A-D) considered in this analysis.
-(b) shows the undeformed experimental setup for the different designs.
-(c) shoes the post-buckling deformed state for each of the designs.
-(d) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.}\label{Fig2}
+	\caption{\textbf{Experimental Results.} (a)-(d) shows the different designs (\emph{Design A-D}) considered in this analysis.
+(e) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.(f) shows the undeformed experimental setup for the different designs.}\label{Fig2}
 \end{figure*}
 
-\begin{figure*}[ht]
+The hexactinellida sponge \textit{Euplectella Asperguillum} species, commonly known as the Venus' Flower Basket, has received significant attention in the engineering and material science communities for its remarkable hierarchical design and robustness across many scales.
+Its fundamental building blocks, called spicules, are constructed using laminated skeletal elements which consist of a central protein-based organic core surrounded by alternating concentric domains of consolidated silica nanoparticles (clear glass) and organic interlayers.
+\cite{weaver2010,aizenberg2005,shulze1887} Remarkably, these spicules are arranged to form a regular and complex square-grid architecture reinforced by a set of double diagonals (a set of two diagonals going in each direction), creating a checkerboard-like pattern of alternating open-closed cells (see \cref{Fig1}).
+Notably, while it has been demonstrated that the laminar architecture of the spicule leads to superior resistance to fracture propagation \cite{miserez2008} and the shape of the spicules enhances the overall strength against buckling \cite{monn2015}, the functionality of the complex double diagonal square-grid macroscale structure is still unknown.
+
+The macro-scale architecture of the \textit{Euplectella Asperguillum} provides a remarkable example of slender and regular lattice-like material constructed by nature.
+Such lattice materials have recently demonstrated to offer novel and unique properties, including being light weight, \cite{schaedler2011,ashby2006} having high energy absorption, \cite{evans2001} and having the ability to control the propagation of waves \cite{phani2006} as well as heat flow.
+\cite{lu1998,ashby1997,evans1998} Generally speaking, the properties and functionality of such materials are  dictated by the node connectivity.
+In particular, a minimum node connectivity of 6 is required for 2D lattice materials to be  stretching-dominated, and therefore,  more weight-efficient for structural applications.
+\cite{deshpande2001} As such, lattice materials with square architecture (which has a node connectivity of 4) are known to be very unstable when loaded along specific directions (as the only shear resistance arises from the joints) \cite{gibson1999} and typically require diagonal bracings to stabilize them.
+
+In this paper, we use the anatomy of the hexactinellid sponge species \textit{Euplectella Aspergilum} skeletal system as inspiration for the design of mechanically robust lattice materials with square architecture.
+First, we use a combination of experiments and numerical analyses to understand the overarching mechanical functions driving the particular and regular alignment of trusses found in the sponge.
+Then, we utilize an optimization algorithm based on the principles introduced by the sponge to determine the  configuration of reinforced square lattice that achieve the highest mechanical performance.
+Therefore, our results indicate that the  principles used by Nature to design the \textit{Euplectella Aspergilum}  may guide the design of lattice architectures structurally more robust than those currently used in modern devices and infrastructure.
+
+To understand  the role of the sponge architecture, we start by comparing its performance to that of other three 2D lattices, all based on a   square architecture with edges of length $L$ (see Supplementary Materials for details).
+Specifically,  we  consider:
+\emph{Design A}  which is inspired by the sponge and comprises a set of double diagonals going in opposite directions (see \cref{Fig2}(a)); \emph{Design B}  which is similar to the sponge design but only contains a single diagonal crossing each of the closed cells (see \cref{Fig2}(b));   \emph{Design C}  which is similar to bracings found in modern structural engineering and contains a crossing set of diagonal beams in every cell (see \cref{Fig2}(c)); and \emph{Design D}  which  has no diagonal reinforcement (see \cref{Fig2}(d)).
+In all our designs we consider struts with rectangular cross section and depth large enough to avoid any out-of-plane deformation.
+Moreover, to ensure a fair mechanical comparison between the four designs, we keep the total mass constant and
+for \emph{Designs A, B}, and \emph{C} we also constrain the  ratio between the mass of the diagonals and non-diagonals struts to be identical (see Supplementary Materials for details).
+%Finally, in all our analyses we constraint the lattices to deform in-plane.
+To this end, in our experiments we consider specimens sufficiently thick for which out-of-plane deformation is negligible and in the the numerical simulations we assume plane-strain conditions.
+
+We started by testing samples of the four considered designs comprising
+a tessellation of $6\times6$ square cells with $L=$ \KB{??} mm.
+The samples were fabricated with a multi-material \mf{@Katia: what do you mean multi-material} 3D printer \mf{(add model @James)} using \KB{name of material} \mf{add info here @James} and compressed uniaxially  using  an Instron Model 5969 with 50kN capacity  (\Cref{Fig2}(e)).
+\KB{switch labels (f) and (e) in Fig 2} %Using this set-up, we tested three replicas of each design that were 3D printed and tested on different days.
+The results reported in \Cref{Fig2}(f)  show that all structures are characterized by an initial linear behavior, a sudden departure from linearity caused by in-plane buckling and catastrophic failure following immediately thereafter.
+Furthermore, it is also important to note that all  designs with diagonal reinforcement (i.e.
+\emph{Designs A-C}) overlay each other in the linear regime.
+Focusing on the initial linear regime, we find that  the different diagonal reinforcement designs considered do not impact the structure's overall stiffness (i.e.
+the initial stiffness of Designs A, B and C is identical).
+Design D, as expected, has a higher stiffness because of its thicker vertical and horizontal element's.
+However, it fails much earlier due to a shorter critical buckling strain.
+On the other hand, \emph{Design A} (the sponge design) outperforms the  considered structures in the amount of load it can bare before the onset of buckling occurs and provides the best mechanical performance without compromising the structural stiffness.
+
+Next, we use Finite Element (FE) simulations conducted using the commercial software ABAQUS/Standard to further understand how the double diagonal system found in the \textit{Euplectella Aspergilum} results in significantly improved mechanical performance.
+All our models are constructed using  Timoshenko beam elements (Abaqus element type \KB{B??}) and the material's response is captured using an isotropic linear elastic materials.
+We start by simulating the response of the  6 by 6 specimens tested in our experiments and find  very good agreement  between our experiments and numerical model (see Fig.
+\KB{??}).
+Having validated our numerical analysis, we next use them to investigate \KB{what???
+big picture of Fig 3} the To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI).
+For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections.
+In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.
+
+The results presented in \cref{Fig3}(a) show that all of the structures containing diagonal reinforcement behave the same way for varying loading angles.
+This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to the perpendicular aligned beam elements.
+This is further supported by the graph for \emph{Design D}, where all of the material from the diagonal beams is removed and allocated to the non-diagonal elements, giving the structure a higher stiffness at $0^\circ$ and lower stiffness at $45^\circ$.
+Any stiffness at $45^\circ$ in Design D is strictly a result of the bending resistance due to the pinned joints between the non-diagonal beams.
+
+The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements.
+From this figure, we can see that \emph{Design A}, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement.
+At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of \emph{Design C}, which is the commonly used truss systems.
+
+%In order to study the impact of diagonal separation for Design A (sponge design), we perform a parametric study on the buckling behavior of the structure under varying separation distances ($C$).
+Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles.
+{Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$).
+Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.
+
+%To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements.
+To quantitatively describe this mass ratio, we introduce the variable $\lambda$ as the fraction $V_{nd}/V_d$, where $V_{nd}$ describes the collective volume of all non-diagonal elements and $V_d$ is the collective volume of all diagonal elements in the unit cell.
+It is important to note, that for all simulations varying $\lambda$ we maintain the same total mass of the structure, namely $V_{nd}+V_d=\mbox{const.
+} \forall (V_{nd},V_d)$.
+
+\mf{write section on Optimization here}
+
+In summary, the investigation of the \textit{Euplectella Asperguillum's} intricate and regular structure has lead to macro-scale design principles that can be applied to general square lattice materials and structures.
+When compared with other diagonally reinforced square lattice designs, the sponge design has experimentally and numerically proven to provide a superior mechanism for withstanding loads prior to onset of buckling.
+By controlling the number of diagonals, diagonal separation and mass ratio between diagonal and non-diagonal elements, a structure's design can be optimized to withstand buckling load caring capacity.
+
+\section*{Acknowledgments}
+ \noindent This work was partially supported by: NSF-GRFP Fellowship  Grant\# DGE-1144152 (for M.
+C.
+F.), GEM Consortium Fellowship (for M.
+C.
+F.), the Harvard Graduate Prize Fellowship (for M.
+C.
+F.), \mf{add whatever else here}.
+ We would also like to thank James R. Rice, Mark Scott, Joost Vlassak, and Chris Rycroft for insightful discussions.
+
+\begin{figure*}[!htb]
+	\centering
 	\captionsetup{width=0.8\textwidth}
-	\begin{center}
 	\includegraphics[width=0.7\textwidth]{Fig3}
-	\end{center}
 	\caption{\textbf{Structure Mechanical Response} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs.
-The designs considered are depicted below as Design A-D.} \label{Fig3}
+The designs considered are depicted below as \emph{Design A-D.}} \label{Fig3}
 \end{figure*}
 
-\begin{figure*}[ht]
+\begin{figure*}[!htb]
+	\centering
 	\captionsetup{width=0.8\textwidth}
-	\begin{center}
 	\includegraphics[width=0.7\textwidth]{Fig4}
-	\end{center}
 	\caption{\textbf{Critical Buckling Load Optimization Results.} (a) shows the optimal value of critical buckling load for varying number of diagonals.
@@
 
-The design of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}.
-In modern day structures, we see many different iterations of new truss designs with various applications in mind but with what is the same fundamentally simple diagonal reinforcement technique.
-This technique, involving two diagonal trusses (one in each direction) crossing the opening of a square beam cell, limits the transverse motion of the main support beams ultimately giving the structure shear capacity and strength.
-Alternative, methods to strengthen beam structures have been introduced via material \hl{[...],} \mf{needs more information and citations}.
-However, in this study we seek to further strengthen square lattices through inspiration from a particular sponge species called \textit{Euplectella Aspergillum}, commonly known as the Venus' flower basket.
-This particular species, which is part of the \textit{Hexactinellid} class, is primarily made of silica and exhibit a periodic and regular grid structure with an intrinsic and different diagonal reinforcement scheme.
-
-\textit{Hexactinellids}, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m.
-Beyond their fracture propagation inhibiting material composition \citep{weaver2007}, they are perceived to exhibit large structural rigidity and strength against buckling.
-
-Structurally, they are composed of a base square-grid architecture and regular ordering of vertical and horizontal struts that form the skeletal system.
-Furthermore, their base structure is overlaid with double diagonal (two diagonals in each direction) reinforcement struts, which create a checkerboard-like pattern of open-closed cell structure.
-This diagonal reinforcement design is conjectured to give the sponge greater buckling resistance and strength to localized damage then it would experience having a single diagonal reinforcement strut (with equivalent allocation of mass between the diagonal reinforcements.)
-
-The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?}
-
-In this paper, we present a numerical and experimental mechanical analysis of multiple  structures under various loading conditions as well as survey different arrangements within similar design space via parametric studies.
-Furthermore, we incorporate an optimization algorithm on our parameter space to obtain an optimal design to withstand (via critical buckling load) uniaxial loading parallel to the non-diagonal truss members.
-For this analysis, we restrict ourselves to four designs considered (Design A-D) that are shown along the top of \cref{Fig2}.
-As part of our analysis, we study the critical buckling strain and the elastic load carrying capacity of these structures.
-
-Numerically, we compare the linear stiffness and critical buckling strain for each of the different designs using Finite Element Analysis (FEA).
-To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI).
-For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections.
-In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.
-
-The results presented in \cref{Fig3}(a) show that all of the structures containing diagonal reinforcement behave the same way for varying loading angles.
-This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to the perpendicular aligned beam elements.
-This is further supported by the graph for Design D, where all of the material from the diagonal beams is removed and allocated to the non-diagonal elements, giving the structure a higher stiffness at $0^\circ$ and lower stiffness at $45^\circ$.
-Any stiffness at $45^\circ$ in Design D is strictly a result of the bending resistance due to the pinned joints between the non-diagonal beams.
-
-The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements.
-From this figure, we can see that Design A, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement.
-At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of Design C, which is the commonly used truss systems.
-
-For the experimental validation, we conducted various uniaxial compression tests of 3D printed planar beam structures (with circular cross-section) on an Instron 5566.
-The specimen was 3D printed using \hl{xxxx} material with manufacturer reported \hl{xxxx GPa} modulus.
-The printed specimen consisted of a 10x10 cell structure spanning a total of 10x10 inch planar structure.
-The experimental set-up consisted of two \hl{1 inch} thickness acrylic sheets bolted together along the perimeter and spaced from each other by the thickness of the 3D printed specimen.\\\\\hl{[...]} \mf{more to go here when we have experimental setup and data}
-
-In order to study the impact of diagonal separation for Design A (sponge design), we perform a parametric study on the buckling behavior of the structure under varying separation distances ($C$).
-Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles.
-{Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$).
-Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.
-
-To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements.
-To quantitatively describe this mass ratio, we introduce the variable $\lambda$ as the fraction $V_{nd}/V_d$, where $V_{nd}$ describes the collective volume of all non-diagonal elements and $V_d$ is the collective volume of all diagonal elements in the unit cell.
-It is important to note, that for all simulations varying $\lambda$ we maintain the same total mass of the structure, namely $V_{nd}+V_d=\mbox{const.
-} \forall (V_{nd},V_d)$.
-
-%\section*{Acknowledgments}
-
-% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship
-% Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft
-
-\nocite{*}
+\noindent\rule[-1.5ex]{\linewidth}{1pt}
+%\nocite{*}
 \bibliography{refs}
-\bibliographystyle{apalike}
+%\bibliographystyle{apalike}
+%\bibliographystyle{plainnat}
+\bibliographystyle{unsrtnat}
 

### 2018-10-24 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
new version

@@
 \newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
-\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}
+\newcommand{\mf}[1]{\noindent\color{color2}\sethlcolor{cyan}\hl{$\Longrightarrow$#1}\normalcolor}
 
-\title{Supplemental Information:\\ Harnessing  Design Principles from Glass Sponges for Structurally Robust Lattice Structures}
+\title{\textit{Supplemental Information:}\\ Design Principles from Glass Sponges \\for Structurally Robust Lattices}
 
@@
 % \doublespace
-% \linenumbers
+\linenumbers
+
 \section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
@@
 However, in this paper we do not report the effects of these ridges on it's structural performance.
+
+\subsection{Image Segmentation Analysis}
+\mf{@James, add details of image segmentation analysis performed for extracting diameter rations between diagonals and non-diagonals}
 
@@
 \end{equation}
-with $\alpha=$ B and C.
+with $\alpha\in$ B and C.
 
@@
 \end{equation}
-with $\alpha=$ B, C and D.
+with $\alpha\in$ B, C and D.
 Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A,
@@
     \centering
-	\includegraphics[width=0.6\linewidth]{SFig10.png}
+	\includegraphics[width=0.9\linewidth]{SFig10.pdf}
 	\caption{{\bf RVE used for the different designs.} This figure portrays the different RVEs used for the different designs considered.
@@
 \end{figure}
+
+\section{Experimental Setup}
+\mf{add full description of experimental setup and full table of 3D printed material properties; @James: I need to material properties}
 
@@
 However, an equivalent analysis is performed to maximize the critical buckling strain as the target response (\cref{BucklingOptimization}.)
+
+\mf{Add table with optimal parameters results for each N, for both circular and rectangular cross-sections}
 
@@
 \end{figure}
+
+\section{Stress Concentration}
+\mf{add section on stress concentration for plane strain setting}
+
 % Bibliography
-%\nocite{aizenberg2005}
-%\nocite{deshpande2001}
-%\nocite{miserez2008}
-%\nocite{weaver2010}
+\bibliography{refs}
+%\bibliographystyle{apalike}
+% \bibliographystyle{plainnat}
+\bibliographystyle{unsrtnat}
 
-\bibliography{refs}
-\bibliographystyle{apalike}
-% \bibliographystyle{plainnat}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{SFig13.pdf}
+	\caption{{\bf asdf.} asdf}
+	\label{asdfa}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{SFig14.pdf}
+	\caption{{\bf asdf.} asdf}
+	\label{asdfgas}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{SFig15.pdf}
+	\caption{{\bf asdf.} asdf}
+	\label{asdfgdfsgas}
+\end{figure}
 

### 2018-12-05 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
changes

@@
 In particular, a minimum node connectivity of 6 is required for 2D lattice materials to be  stretching-dominated, and therefore,  more weight-efficient for structural applications.
-\cite{deshpande2001} As such, lattice materials with square architecture (which has a node connectivity of 4) are known to be very unstable when loaded along specific directions (as the only shear resistance arises from the joints) \cite{gibson1999} and typically require diagonal bracings to stabilize them.
+\cite{deshpande2001} As such, lattice materials with square architecture (which has a node connectivity of 4) are known to be very unstable when loaded along specific directions (as the only shear resistance arises from the joints) \cite{gibson1999} and typically require diagonal bracings to stabilize them.\cite{phani2017}
 
@@
 We started by testing samples of the four considered designs comprising
-a tessellation of $6\times6$ square cells with $L=$ \KB{??} mm.
-The samples were fabricated with a multi-material \mf{@Katia: what do you mean multi-material} 3D printer \mf{(add model @James)} using \KB{name of material} \mf{add info here @James} and compressed uniaxially  using  an Instron Model 5969 with 50kN capacity  (\Cref{Fig2}(e)).
+a tessellation of $6\times6$ square cells with $L=1.91$ cm.
+The samples were fabricated with a multi-material 3D printer \mf{(add model @James)} using \KB{name of material} \mf{add info here @James} and compressed uniaxially  using  an Instron Model 5969 with 50kN capacity  (\Cref{Fig2}(e)).
 \KB{switch labels (f) and (e) in Fig 2} %Using this set-up, we tested three replicas of each design that were 3D printed and tested on different days.

### 2018-12-05 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
changes

@@
     \centering
-    \includegraphics[width=0.4\linewidth]{SFig2.png}
+    \includegraphics[width=0.5\linewidth]{SFig2}
     \caption{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system.
@@
 In an effort to identify lattice configurations resulting in large critical loads, we used a Python implementation of the Covariance Matrix Adaptation Evolution Strategy (CMA-ES) \citep{hansen2003}.
-CMA-ES is an evolutionary
-algorithm that is used to solve optimization/inverse problems by iteratively solving several forward problems to adjust a covariance matrix of the solution.
+CMA-ES is an evolutionary algorithm that is used to solve optimization/inverse problems by iteratively solving several forward problems to adjust a covariance matrix of the solution.
 CMA-ES is a derivative free algorithm, well suited for optimization problems of high dimensionality and non-linear parameter topology.
@@
 \end{itemize}
-resulting in a lattice structure with the large critical load.
+resulting in a lattice structure with the largest critical load.
 For such optimization problem, the number of optimization variables increases with the number of diagonals incorporated in the model(i.e.
@@
 
-\mf{Add table with optimal parameters results for each N, for both circular and rectangular cross-sections}
+\begin{table}[!htb]
+\centering
+    \begin{tabular}{c||c|c|c|c|c}
+          & $\lambda$ & $S_1$ &$S_2$ &$S_3$&$S_4$ \\ \hline\hline
+         $N=1$ &  3.1454 &  &  &  &  \\ \hline
+         $N=2$ &  0.5614 &  0.3390 &  &  &  \\ \hline
+         $N=3$ &  1.4784 &  0.2440 &  &  &  \\ \hline
+         $N=4$ &  1.0151 &  0.0989 &  0.3358 &  &  \\ \hline
+         $N=5$ &  0.9509 &  0.1733 &  0.3260 &  &  \\ \hline
+         $N=6$ &  0.2009 &  0.2628 &  0.5827 &  0.8881 &  \\ \hline
+         $N=7$ &  0.2962 &  0.4197 &  0.6917 &  0.9126 &  \\ \hline
+         $N=8$ &  0.3263 &  0.1553 &  0.5533 &  0.7095 &  0.9113 \\
+    \end{tabular}
+\caption{Buckling Load Optimization}
+\label{table1}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+    \begin{tabular}{c||c|c|c|c|c}
+          & $\lambda$ & $S_1$ &$S_2$ &$S_3$&$S_4$ \\ \hline\hline
+         $N=1$ &  0.8908 &  &  &  &  \\ \hline
+         $N=2$ &  0.4157 &  0.3233 &  &  &  \\ \hline
+         $N=3$ &  0.3402 &  0.3268 &  &  &  \\ \hline
+         $N=4$ &  0.2271 &  0.6103 &  0.7923 &  &  \\ \hline
+         $N=5$ &  0.2772 &  0.2511 &  0.4076 &  &  \\ \hline
+         $N=6$ &  0.1764 &  0.2785 &  0.5743 &  0.8675 &  \\ \hline
+         $N=7$ &  0.1810 &  0.2899 &  0.5711 &  0.8443 &  \\ \hline
+         $N=8$ &  0.1023 &  0.1459 &  0.3317 &  0.5561 &  0.7843 \\
+    \end{tabular}
+\caption{Buckling Strain Optimization}
+\label{table2}
+\end{table}
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{SFig16.pdf}
+	\caption{{\bf Stress Concentration.} Highest Misses Stress for different designs: Design A -- $3.362\times 10^{-2}$, Design B -- $4.362\times 10^{-2}$, Design C -- $4.165\times 10^{-2}$, Design D -- $3.550\times 10^{-2}$.}
+	\label{asdfgdfsgasasdf}
+\end{figure}
+
 \end{document}

### 2018-12-07 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
added figs in supplemental
.

@@
 
-To understand  the role of the sponge architecture, we start by comparing its performance to that of other three 2D lattices, all based on a   square architecture with edges of length $L$ (see Supplementary Materials for details).
+To understand the role of the sponge architecture, we start by comparing its performance to that of additional three 2D lattices, all based on a square architecture with edge lengths $L$ (see Supplementary Materials for details).
 Specifically,  we  consider:
@@
 
-We started by testing samples of the four considered designs comprising
-a tessellation of $6\times6$ square cells with $L=1.91$ cm.
+We began our analysis by testing samples of the four considered designs comprising a tessellation of $6\times6$ square cells with $L=1.91$ cm.
 The samples were fabricated with a multi-material 3D printer \mf{(add model @James)} using \KB{name of material} \mf{add info here @James} and compressed uniaxially  using  an Instron Model 5969 with 50kN capacity  (\Cref{Fig2}(e)).

### 2018-12-09 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

@@
 	\caption{\textbf{Experimental Results.} (a)-(d) shows the different designs (\emph{Design A-D}) considered in this analysis.
-(e) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.(f) shows the undeformed experimental setup for the different designs.}\label{Fig2}
+(e) shows the undeformed experimental setup for the different designs.
+(f) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.}\label{Fig2}
 \end{figure*}
@@
 We began our analysis by testing samples of the four considered designs comprising a tessellation of $6\times6$ square cells with $L=1.91$ cm.
-The samples were fabricated with a multi-material 3D printer \mf{(add model @James)} using \KB{name of material} \mf{add info here @James} and compressed uniaxially  using  an Instron Model 5969 with 50kN capacity  (\Cref{Fig2}(e)).
-\KB{switch labels (f) and (e) in Fig 2} %Using this set-up, we tested three replicas of each design that were 3D printed and tested on different days.
+The samples were fabricated with a multi-material 3D printer \mf{(add model @James)} using \KB{name of material @James} and compressed uniaxially using an Instron Model 5969 with 50kN capacity (\Cref{Fig2}(e)).
+For each design, we 3D print three separate specimens that are then reported in \cref{Fig2}(f).
 The results reported in \Cref{Fig2}(f) show that all structures are characterized by an initial linear behavior, a sudden departure from linearity caused by in-plane buckling and catastrophic failure following immediately thereafter.
@@
 Focusing on the initial linear regime, we find that  the different diagonal reinforcement designs considered do not impact the structure's overall stiffness (i.e.
-the initial stiffness of Designs A, B and C is identical).
-Design D, as expected, has a higher stiffness because of its thicker vertical and horizontal element's.
+the initial stiffness of Designs A, B and C are identical).
+Design D, as expected, has a higher stiffness because of its thicker vertical and horizontal elements.
 However, it fails much earlier due to a shorter critical buckling strain.
-On the other hand, \emph{Design A} (the sponge design) outperforms the  considered structures in the amount of load it can bare before the onset of buckling occurs and provides the best mechanical performance without compromising the structural stiffness.
+On the other hand, \emph{Design A} (the sponge design) outperforms the considered structures in the amount of load it can bare before the onset of buckling occurs and provides the best mechanical performance without compromising the structural stiffness while using the same total amount of material.
 
-Next, we use Finite Element (FE) simulations conducted using the commercial software ABAQUS/Standard to further understand how the double diagonal system found in the \textit{Euplectella Aspergilum} results in significantly improved mechanical performance.
-All our models are constructed using  Timoshenko beam elements (Abaqus element type \KB{B??}) and the material's response is captured using an isotropic linear elastic materials.
-We start by simulating the response of the  6 by 6 specimens tested in our experiments and find  very good agreement  between our experiments and numerical model (see Fig.
-\KB{??}).
-Having validated our numerical analysis, we next use them to investigate \KB{what???
-big picture of Fig 3} the To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI).
-For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections.
-In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.
+Next, we use Finite Element (FE) simulations conducted with the widely used commercial software ABAQUS/Standard to further understand how the double diagonal system found in the \textit{Euplectella Aspergilum} results in significantly improved mechanical performance.
+All our models are constructed using Timoshenko beam elements (ABAQUS element type B22) and the material's response is captured using an isotropic linear elastic materials.
+We start by simulating the response of the 6 by 6 specimens tested in our experiments and find very good agreement between our experiments and numerical model (see \cref{Fig2}(f)).
+Having validated our numerical analysis with experimental results, we next use the numerical model to investigate the first-order behavior of the structure (neglecting all edge effects by the assumption of an infinite periodic structure).
+To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI).
+In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for the designs containing diagonal reinforcement.
 
 The results presented in \cref{Fig3}(a) show that all of the structures containing diagonal reinforcement behave the same way for varying loading angles.
-This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to the perpendicular aligned beam elements.
+This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to elements aligned parallel to the angle of loading.
 This is further supported by the graph for \emph{Design D}, where all of the material from the diagonal beams is removed and allocated to the non-diagonal elements, giving the structure a higher stiffness at $0^\circ$ and lower stiffness at $45^\circ$.
@@
 
-%In order to study the impact of diagonal separation for Design A (sponge design), we perform a parametric study on the buckling behavior of the structure under varying separation distances ($C$).
-Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles.
-{Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$).
-Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.
-
-%To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements.
-To quantitatively describe this mass ratio, we introduce the variable $\lambda$ as the fraction $V_{nd}/V_d$, where $V_{nd}$ describes the collective volume of all non-diagonal elements and $V_d$ is the collective volume of all diagonal elements in the unit cell.
-It is important to note, that for all simulations varying $\lambda$ we maintain the same total mass of the structure, namely $V_{nd}+V_d=\mbox{const.
-} \forall (V_{nd},V_d)$.
+Multiplying the results of \cref{Fig3}(a) together with \cref{Fig3}(b) provides a measure of the normalized critical buckling load as seen in \cref{Fig3}(c).
+From these results, we clearly see that the geometric benefit attributed to \emph{Design A} clearly propagates through all loading angles.
+The benefit presented, is substantial near $0^\circ$ loading angles from all other designs.
+For loading angles near $\approx 45 ^\circ$, \emph{Design A} still outperforms \emph{Design B} and \emph{Design C}.
+However, it is important to note that \emph{Design D} appears to have a benefit over \emph{Design A} in this regime.
+This benefit, however, is attributed to numerical constraints imposed by the periodic boundary conditions.
+When creating a finite size structure and omit and peridic boundary conditions from our model, we see the benefit of \emph{Design D} over \emph{Design A} vanish and \emph{Design A} outperforms over all other considered designs for all loading angles.
+This further benefit for \emph{Design A} is attributed to it's insensitivity to edge effects.
+Furthermore, from \cref{Fig3}(e), we can see that \emph{Design A}, \emph{Design B}, \emph{Design C} have periodic buckling modes that are note dependent on the domain size, whereas for \emph{Design D}, the buckling mode strongly depends on the domain size (find further analysis on varying loading angles in the SI.)
 
